---
title: "STAR-Fusion_v1.7_TCGA_n_GTEx"
author: "Brian Haas"
date: "2/9/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```


```{r}
data = read.table(gzfile("../data/TCGA_n_GTEx.STAR-Fusion.v1.7.fusion_preds.decorated.tsv.gz"), header=T, sep="\t", stringsAsFactors = F)

data = data %>% mutate(dataclassTN = paste0(data_class, "-", tumor_or_normal))

#Count numbers of samples before basic filtering

data %>% select(dataclassTN, sample_name) %>% unique() %>% ungroup() %>% group_by(dataclassTN) %>% tally()

## some initial cleanup
data = data %>% 
  filter(! grepl("chrM", Annots)) %>%
  filter(! grepl("HLA", fusion_name)) %>%
  filter(! (grepl("^IG[HKVL]", LeftGene) & grepl("^IG[HKVL]", RightGene)))

## only consider the primary isoform.
data = data %>% filter(is_primary)


# require reference splicing
data = data %>% filter(SpliceType == "ONLY_REF_SPLICE")


#Count numbers of samples post basic filtering

data %>% select(dataclassTN, sample_name) %>% unique() %>% ungroup() %>% group_by(dataclassTN) %>% tally()

```




```{r}

get_FPM_vs_min_FFPM_via_tissue_type = function (data, min_ffpm_vals = seq(0, 0.3, 0.05)) {
    
    fusion_counts_vs_min_ffpm = NULL
    
    
    # ignore READTHRU fusions here
    
    samples_n_tissues = data %>%   
      filter(structure_type != "READTHRU") %>% 
      select(data_class, dataclassTN, tissue_type, sample_name) %>% unique()
    
    
    for (min_ffpm in min_ffpm_vals) {
        message(min_ffpm)
        
        data_filt = data %>% filter(FFPM >= min_ffpm)
        
        counts_per_filt_sample = data_filt %>% select(data_class, dataclassTN, tissue_type, sample_name, fusion_name) %>% 
            unique() %>% group_by(data_class, dataclassTN, tissue_type, sample_name) %>% tally(name='fusion_count')
        
        counts_per_filt_sample = left_join(samples_n_tissues, counts_per_filt_sample, 
                                           by=c('data_class', 'dataclassTN', 'tissue_type', 'sample_name')) %>%
        mutate(fusion_count = ifelse(is.na(fusion_count), 0, fusion_count))
        
        
        
        counts_per_filt_sample$min_ffpm = min_ffpm
        
        fusion_counts_vs_min_ffpm = rbind(fusion_counts_vs_min_ffpm, counts_per_filt_sample)
        
    }
    
    return(fusion_counts_vs_min_ffpm)
}


```


```{r}



tissue_fusion_counts_vs_min_ffpm = get_FPM_vs_min_FFPM_via_tissue_type(data)


```


Compute median fusion count for each tissue type

```{r}
median_tissue_fusion_counts = tissue_fusion_counts_vs_min_ffpm  %>% 
    group_by(data_class, dataclassTN, tissue_type, min_ffpm) %>% 
    summarize(median_fusion_count = median(fusion_count)) %>% ungroup()

median_tissue_fusion_counts 
```

Counts of GTEx-normal fusions vs. min FFPM

```{r}
median_tissue_fusion_counts %>% filter(dataclassTN == "GTEx-normal") %>% 
    ggplot(aes(x=min_ffpm, y=median_fusion_count, color=tissue_type)) + geom_point() + geom_line()

```

Counts of TCGA-tumor fusions vs. min FFPM

```{r}
median_tissue_fusion_counts %>% filter(dataclassTN == "TCGA-tumor") %>% 
    ggplot(aes(x=min_ffpm, y=median_fusion_count, color=tissue_type)) + geom_point() + geom_line()

```


Jitter plots for counts of fusions per tissue type according to data class (ie. tcga-tumor, tcga-normal, and gtex-normal)
```{r}


min_ffpms = median_tissue_fusion_counts %>% select(min_ffpm) %>% unique() %>% arrange(min_ffpm) %>% pull(min_ffpm)

for (min_ffpm_use in min_ffpms) {
    
    
    p = median_tissue_fusion_counts %>% filter(min_ffpm == min_ffpm_use) %>%
        ggplot(aes(x = dataclassTN, y=median_fusion_count)) + 
        geom_point(position = position_jitter(w = 0.1, h = 0)) +
        ggtitle(paste("median fusion counts/sample with min FFPM=", min_ffpm_use))
    
       
    plot(p)
    
}
```


Plot TCGA tumor/normal median fusion counts according to matched sample types.

```{r}

tcga_tissues_with_matched_normals = median_tissue_fusion_counts %>% filter(dataclassTN == "TCGA-normal") %>% 
    select(tissue_type) %>% unique() %>% pull(tissue_type)


for (min_ffpm_use in min_ffpms) {
    
  # statistical analysis of medians
  tumor_normal_paired_medians = median_tissue_fusion_counts %>% 
    filter(data_class == "TCGA" & tissue_type %in% tcga_tissues_with_matched_normals & min_ffpm == min_ffpm_use) %>%
    arrange(tissue_type) %>% spread(key=dataclassTN, value=median_fusion_count)
  
   normal_medians = tumor_normal_paired_medians %>% select(`TCGA-normal`) %>% pull(`TCGA-normal`)
   tumor_medians = tumor_normal_paired_medians %>% select(`TCGA-tumor`) %>% pull(`TCGA-tumor`)
   t = t.test(tumor_medians, normal_medians, paired = T)
   print(min_ffpm_use)
   print(t)
   
   p = median_tissue_fusion_counts %>% filter(data_class == "TCGA" & tissue_type %in% tcga_tissues_with_matched_normals) %>%
    filter(min_ffpm == min_ffpm_use) %>%
    ggplot(aes(x=tissue_type, y=median_fusion_count, fill=dataclassTN)) + geom_col(position='dodge') + 
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(1))) +
       ggtitle(paste("Median TCGA fusion counts at min FFPM threshold: ", min_ffpm_use))
   
   plot(p)
   
}

```


```{r}


for (min_ffpm_use in min_ffpms) {

    p =  median_tissue_fusion_counts %>% 
        filter(min_ffpm == min_ffpm_use) %>% 
        group_by(dataclassTN, min_ffpm) %>% 
        summarize(median_of_medians = median(median_fusion_count)) %>%
        ggplot(aes(x=dataclassTN, y=median_of_medians)) + geom_col() + 
        ggtitle(paste("Median of median fusion counts, min ffpm = ", min_ffpm_use))
    
    plot(p)

}

```


# Examine distribution of fusion expression levels

```{r}

data2 = data %>% mutate(dataclass2 = dataclassTN) %>% 
    mutate(dataclass2 = ifelse(grepl("Cosmic",Annots), "Cosmic", dataclass2))

cosmic_fusion_FFPM = data2 %>% filter(dataclass2 == "Cosmic") %>% pull(FFPM)
GTEx_normal_FFPM = data2 %>% filter(dataclassTN == "GTEx-normal") %>% pull(FFPM)
TCGA_tumor_FFPM = data2 %>% filter(dataclassTN == "TCGA-tumor") %>% pull(FFPM)
TCGA_normal_FFPM = data2 %>% filter(dataclassTN == "TCGA-normal") %>% pull(FFPM)

df_FFPM = bind_rows( data.frame(dataclass="Cosmic", FFPM=cosmic_fusion_FFPM),
                     data.frame(dataclass="GTEx-normal", FFPM=GTEx_normal_FFPM),
                     data.frame(dataclass="TCGA-tumor", FFPM=TCGA_tumor_FFPM),
                     data.frame(dataclass="TCGA-normal", FFPM=TCGA_normal_FFPM) )
                     
df_FFPM %>% mutate(FFPM_max1 = pmin(1, FFPM)) %>% ggplot(aes(x=FFPM_max1, color=dataclass)) + geom_density(alpha=0.2) + xlim(0, 1)

```

Compare ECDFs for the above.
```{r}

cosmic_fusion_FFPM_ecdf = ecdf(cosmic_fusion_FFPM)
GTEx_normal_FFPM_ecdf = ecdf(GTEx_normal_FFPM)
TCGA_tumor_FFPM_ecdf = ecdf(TCGA_tumor_FFPM)
TCGA_normal_FFPM_ecdf = ecdf(TCGA_normal_FFPM)

ecdf_range = seq(0, 3, 0.01)

df_ecdf = bind_rows( data.frame(dataclass="Cosmic", minFFPM=ecdf_range, cum_frac=cosmic_fusion_FFPM_ecdf(ecdf_range)),
                     data.frame(dataclass="GTEx-normal", minFFPM=ecdf_range, cum_frac=GTEx_normal_FFPM_ecdf(ecdf_range)),
                     data.frame(dataclass="TCGA-tumor", minFFPM=ecdf_range, cum_frac=TCGA_tumor_FFPM_ecdf(ecdf_range)),
                     data.frame(dataclass="TCGA-normal", minFFPM=ecdf_range, cum_frac=TCGA_normal_FFPM_ecdf(ecdf_range)) )

df_ecdf %>% ggplot(aes(x=minFFPM, y=cum_frac, color=dataclass)) + geom_point()

```

```{r}
message("GTEx ecdf at 0.1 ffpm: ", GTEx_normal_FFPM_ecdf(0.1))

message("cosmic fusions ecdf at 0.1 ffpm: ", cosmic_fusion_FFPM_ecdf(0.1))

```

```{r}

num_gtex_ge = length(GTEx_normal_FFPM[GTEx_normal_FFPM >= 0.1])
num_gtex_lt = length(GTEx_normal_FFPM[GTEx_normal_FFPM < 0.1])

num_cosmic_ge = length(cosmic_fusion_FFPM[cosmic_fusion_FFPM >= 0.1])
num_cosmic_lt = length(cosmic_fusion_FFPM[cosmic_fusion_FFPM < 0.1])

m = matrix(c(num_gtex_ge, num_gtex_lt, num_cosmic_ge, num_cosmic_lt), nrow=2, byrow = T)
fisher.test(m)


```


## Examine the COSMIC fusions:

```{r}

cosmic_fusion_occurrence = data2 %>% filter(data_class == "GTEx" | tumor_or_normal == "tumor") %>% 
    filter(dataclass2 == "Cosmic") %>% 
    select(fusion_name, sample_name, data_class, tissue_type) %>% 
    unique() %>% 
    mutate(sample_tissue = paste(data_class, tissue_type)) %>% 
    select (fusion_name, sample_tissue) %>% 
    group_by(fusion_name, sample_tissue) %>% 
    tally()


sample_tissue_counts = data2 %>% filter(data_class == "GTEx" | tumor_or_normal == "tumor") %>% 
    select(sample_name, data_class, tissue_type) %>% 
    mutate(sample_tissue = paste(data_class, tissue_type)) %>% 
    select(sample_name, sample_tissue) %>% 
    unique() %>% 
    group_by(sample_tissue) %>% 
    tally()

sample_tissue_counts = sample_tissue_counts %>% rename(sample_tissue_count = n)

cosmic_fusion_occurrence = left_join(x=cosmic_fusion_occurrence, y=sample_tissue_counts, by='sample_tissue')

cosmic_fusion_occurrence = cosmic_fusion_occurrence %>% mutate(sample_pct = n / sample_tissue_count * 100)

ranked_fusions = cosmic_fusion_occurrence %>% 
    group_by(fusion_name) %>% 
    summarize(sumpct= sum(sample_pct)) %>% 
    arrange(desc(sumpct)) %>% 
    select(fusion_name)
 
cosmic_fusion_occurrence %>% 
    ggplot(aes(x=factor(fusion_name, levels=ranked_fusions$fusion_name), y=sample_pct, fill=sample_tissue)) + 
    geom_col() + 
    theme(axis.text.x=element_text(angle = 90, hjust = 0, size=rel(0.5))) 
```



Examine distribution of FFPM for cosmic fusions

```{r}


data2 %>% filter(dataclass2 == "Cosmic") %>% 
    ggplot(aes(x=reorder(fusion_name, FFPM, FUN = median), y=FFPM)) + 
    geom_boxplot() +
     theme(axis.text.x=element_text(angle = 90, hjust = 0, size=rel(0.5))) + 
    scale_y_continuous(trans='log2') + 
    ggtitle("Cosmic Fusion Expression Distribution")

```



Examine distribution of structural categories according to min FFPM and data class

```{r}

min_ffpm_vals = seq(0, 0.3, 0.05)

category_pcts_vs_min_ffpm = NULL

for (min_FFPM in min_ffpm_vals) {
  
  
  
  category_pcts = data %>% filter(structure_type != "OTHER") %>% 
    filter(FFPM >= min_FFPM) %>%
    group_by(dataclassTN, structure_type) %>% 
    tally() %>% 
    mutate(pct=prop.table(n)) 
  
  cosmic_pcts = data %>% filter(cosmic) %>% 
    filter(FFPM >= min_FFPM) %>%
    mutate(dataclassTN = "COSMIC") %>%
    group_by(dataclassTN, structure_type) %>% 
    tally() %>% 
    mutate(pct=prop.table(n))
  
  
  category_pcts = bind_rows(category_pcts, cosmic_pcts)
  
  category_pct_plot = category_pcts %>% ggplot(aes(x=dataclassTN, y=pct, fill=structure_type) ) + geom_col() +
    ggtitle(paste("minFFPM = ", min_FFPM))
  
  category_pcts$min_FFPM = min_FFPM
  category_pcts_vs_min_ffpm = bind_rows(category_pcts_vs_min_ffpm, category_pcts)
  
  plot(category_pct_plot)
}

```


```{r}

category_pcts_vs_min_ffpm  %>% ggplot(aes(x=min_FFPM, y=pct, shape=structure_type, color=structure_type)) + geom_line() + geom_point() + facet_wrap(~dataclassTN) + ylim(c(0,0.8))

```


# misc queries

```{r}
# restructure statistics to take into account all tissue types including tcga normals.

# first get counts of all samples

dataclassTN_tissue_counts = data %>% select(dataclassTN, tissue_type, sample_name) %>%
  unique() %>% 
  group_by(dataclassTN, tissue_type) %>% 
  tally(name='sample_count')


dataclassTN_tissue_fusion_sample_occurrence_counts = data %>% select(dataclassTN, tissue_type, sample_name, fusion_name) %>% unique() %>% group_by(dataclassTN, tissue_type, fusion_name) %>% tally(name='fusion_tissue_count')


dataclassTN_tissue_fusion_stats = left_join(dataclassTN_tissue_fusion_sample_occurrence_counts,
                                            dataclassTN_tissue_counts, 
                                            by=c('dataclassTN', 'tissue_type'))

dataclassTN_tissue_fusion_stats = dataclassTN_tissue_fusion_stats %>% mutate(frac_tissue_type = fusion_tissue_count/sample_count)

```



```{r}

dataclassTN_tissue_fusion_stats  %>% filter(fusion_name == "TMPRSS2--ERG")

```




```{r}

dataclassTN_tissue_fusion_stats  %>% filter(fusion_name == "FGFR3--TACC3") %>% 
  arrange(desc(frac_tissue_type))


```
