---
title: "STAR-Fusion_v1.7_TCGA_n_GTEx"
author: "Brian Haas"
date: "2/9/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(extrafont)
```

# Examining preponderance of fusions in tumors and normal tissues

Inputs: STAR-Fusion predictions across TCGA and GTEx

```{r}
data = read.table(gzfile("../data/TCGA_n_GTEx.STAR-Fusion.v1.7.fusion_preds.decorated.tsv.gz"), header=T, sep="\t", stringsAsFactors = F)

data = data %>% mutate(dataclassTN = paste0(data_class, "-", tumor_or_normal))

#Count numbers of samples before basic filtering

data %>% select(dataclassTN, sample_name) %>% unique() %>% ungroup() %>% group_by(dataclassTN) %>% tally()

```

Perform some initial filtering, removing mitochondrial fusions, HLA and immunoglobulin fusions so that they don't confound aspects of our analysis.

```{r}

## some initial cleanup
data = data %>% 
  filter(! grepl("chrM", Annots)) %>%
  filter(! grepl("HLA", fusion_name)) %>%
  filter(! (grepl("^IG[HKVL]", LeftGene) & grepl("^IG[HKVL]", RightGene)))

```


To focus on fusion occurrences, we restrict our analysis to the primary isoforms and exclude the more lowly expressed alternative fusion isoforms where they exist.

Also, we restrict to those isoforms that have breakpoints consistent with splice sites of reference gene structures to minimize the impact of RT-artifacts contributing to this analysis.

```{r}


## only consider the primary isoform.
data = data %>% filter(is_primary)


# require reference splicing
data = data %>% filter(SpliceType == "ONLY_REF_SPLICE")


#Count numbers of samples post basic filtering

data %>% select(dataclassTN, sample_name) %>% unique() %>% ungroup() %>% group_by(dataclassTN) %>% tally()

```


Compute fusion counts per sample and median fusion counts per tissue type according to minimum fusion expression (FFPM) value:


```{r}

get_fusion_counts_vs_min_FFPM_via_tissue_type = function (data, min_ffpm_vals = seq(0, 0.3, 0.025)) {
    
    fusion_counts_vs_min_ffpm = NULL
    
    
    # ignore READTHRU fusions here
    
    samples_n_tissues = data %>%   
      filter(structure_type != "READTHRU") %>% 
      select(data_class, dataclassTN, tissue_type, sample_name) %>% unique()
    
    
    for (min_ffpm in min_ffpm_vals) {
        message(min_ffpm)
        
        data_filt = data %>% filter(FFPM >= min_ffpm)
        
        counts_per_filt_sample = data_filt %>% select(data_class, dataclassTN, tissue_type, sample_name, fusion_name) %>% 
            unique() %>% group_by(data_class, dataclassTN, tissue_type, sample_name) %>% tally(name='fusion_count')
        
        counts_per_filt_sample = left_join(samples_n_tissues, counts_per_filt_sample, 
                                           by=c('data_class', 'dataclassTN', 'tissue_type', 'sample_name')) %>%
        mutate(fusion_count = ifelse(is.na(fusion_count), 0, fusion_count))
        
        
        
        counts_per_filt_sample$min_ffpm = min_ffpm
        
        fusion_counts_vs_min_ffpm = rbind(fusion_counts_vs_min_ffpm, counts_per_filt_sample)
        
    }
    
    return(fusion_counts_vs_min_ffpm)
}




tissue_fusion_counts_vs_min_ffpm = get_fusion_counts_vs_min_FFPM_via_tissue_type(data)


median_tissue_fusion_counts = tissue_fusion_counts_vs_min_ffpm  %>% 
    group_by(data_class, dataclassTN, tissue_type, min_ffpm) %>% 
    summarize(median_fusion_count = median(fusion_count)) %>% ungroup()

median_tissue_fusion_counts 
```

Counts of GTEx-normal fusions vs. min FFPM

```{r}
median_tissue_fusion_counts %>% filter(dataclassTN == "GTEx-normal") %>% 
    ggplot(aes(x=min_ffpm, y=median_fusion_count, color=tissue_type)) + geom_point() + geom_line()

```

Counts of TCGA-tumor fusions vs. min FFPM

```{r}
median_tissue_fusion_counts %>% filter(dataclassTN == "TCGA-tumor") %>% 
    ggplot(aes(x=min_ffpm, y=median_fusion_count, color=tissue_type)) + geom_point() + geom_line()

```


Jitter plots for counts of fusions per tissue type according to data class (ie. tcga-tumor, tcga-normal, and gtex-normal)

- Used to generate Figure S4a

```{r}


min_ffpms = median_tissue_fusion_counts %>% select(min_ffpm) %>% unique() %>% arrange(min_ffpm) %>% pull(min_ffpm)

median_count_stats = NULL

for (min_ffpm_use in min_ffpms) {
    
    print(paste0("Min FFPM: ", min_ffpm_use))
    
    median_counts_min_ffpm  = median_tissue_fusion_counts %>% filter(min_ffpm == min_ffpm_use)
    
    p = median_counts_min_ffpm %>%
        ggplot(aes(x = dataclassTN, y=median_fusion_count)) + 
        geom_point(position = position_jitter(w = 0.1, h = 0)) +
        ggtitle(paste("median fusion counts/sample with min FFPM=", min_ffpm_use))
    
    
    GTEx_normal_counts = median_counts_min_ffpm %>% filter(dataclassTN == "GTEx-normal") %>% pull(median_fusion_count)
    
    TCGA_normal_counts = median_counts_min_ffpm %>% filter(dataclassTN == "TCGA-normal") %>% pull(median_fusion_count)
    
    TCGA_tumor_counts = median_counts_min_ffpm %>% filter(dataclassTN == "TCGA-tumor") %>% pull(median_fusion_count)
    
    print("GTEx-normal vs. TCGA-normal")
    t = t.test(GTEx_normal_counts, TCGA_normal_counts)
    print(t)
    median_count_stats = bind_rows(median_count_stats, data.frame(min_ffpm=min_ffpm_use, comparison = "GTEx-normal vs. TCGA-normal", tp_value = t$p.value))
    
    print("GTEx-normal vs. TCGA-tumor")
    t = t.test(GTEx_normal_counts, TCGA_tumor_counts)
    print(t)
    median_count_stats = bind_rows(median_count_stats, data.frame(min_ffpm=min_ffpm_use, comparison = "GTEx-normal vs. TCGA-tumor", tp_value = t$p.value ))
    
    print("TCGA-normal vs. TCGA-tumor")
    t = t.test(TCGA_normal_counts, TCGA_tumor_counts)
    print(t)
    median_count_stats = bind_rows(median_count_stats, data.frame(min_ffpm=min_ffpm_use, comparison = "TCGA-normal vs. TCGA-tumor", tp_value = t$p.value  ))
    
    plot(p)
    
}
```


Examine the statistical significance of differences between the sets of median fusion counts

```{r}

median_count_stats = median_count_stats %>% mutate(tnegLogP = -1 * log(tp_value))
median_count_stats %>% ggplot(aes(x=min_ffpm, y=tnegLogP, color=comparison)) + geom_point() + geom_line() + ggtitle("t-test -Log(p)")

median_count_stats 

```

We generally don't see significant differences between GTEx normal and TCGA tumors with unfiltered fusion predictions and when filtering according to min 0.1 FFPM, and while there are peculiar significant differences observed at 0>min_ffpm>0.1 FFPM filtering, overall we have little evidence for differences and they dissipate with more reasonable filtering criteria (min 0.1 FFPM).  It is encouraging that we see increasing statistical significance between TCGA-tumor and TCGA-normal median fusion counts as we increase the fusion expression threshold. This is explored further below with more precise testing of paired samples according to tissue types.


As shown below, the median counts of the medians of fusions per sample type are 4 to 5 fusions per sample type when filtering at 0.1 FFPM.
```{r}

median_tissue_fusion_counts %>% filter(min_ffpm == 0.1) %>% group_by(dataclassTN) %>% summarize(median(median_fusion_count))

```

Pancreas is unusual, with a median count of 27 fusions at 0.1 min FFPM threshold. We investigate fusions in pancreas more extensively later, and - spoiler alert - we basically find that many of these pancreas fusions result from very highly expressed fusion partners, and while some look to possibly result from RT-artifacts, many of them generate in-frame fusions and could result from trans-splicing, so highly suspicious but potentially functional products worth exploring further.

```{r}

median_tissue_fusion_counts %>% filter(min_ffpm == 0.1) %>% filter(tissue_type == "Pancreas")

```



Below we examine paired tumor/normal samples from TCGA to see whether we detect differences in fusion counts, as we would expect.

Here we plot TCGA tumor/normal median fusion counts according to matched sample types and examine the effect of filtering based on min FFPM.

- used for Figure S4b


```{r}


tcga_tissues_with_matched_normals = median_tissue_fusion_counts %>% filter(dataclassTN == "TCGA-normal") %>% 
    select(tissue_type) %>% unique() %>% pull(tissue_type)


source("../Rlib/geom_split_violin.R")

tcga_sample_info = data %>% filter(data_class == "TCGA") %>% select(tissue_type, sample_name) %>% unique()
tcga_sample_info$patient = str_replace(tcga_sample_info$sample_name, "-(NT|TP)$", "")
patients_with_both_tumor_normal_samples = tcga_sample_info %>% group_by(patient) %>% tally() %>% filter(n>1) %>% pull(patient)

tissue_fusion_counts_vs_min_ffpm$patient = str_replace(tissue_fusion_counts_vs_min_ffpm$sample_name, "-(NT|TP)$", "")

tumor_normal_compare_pvalues = NULL

for (min_ffpm_use in min_ffpms) {
    
  # statistical analysis of paired counts
  tumor_normal_paired_fusion_counts = tissue_fusion_counts_vs_min_ffpm %>% 
    filter(patient %in% patients_with_both_tumor_normal_samples) %>%
    filter(data_class == "TCGA" & tissue_type %in% tcga_tissues_with_matched_normals & min_ffpm == min_ffpm_use) %>%
    select(data_class, dataclassTN, patient, tissue_type, fusion_count) %>%
    arrange(tissue_type) 

  
   tissue_types = tumor_normal_paired_fusion_counts %>% select(tissue_type) %>% unique() %>% pull(tissue_type)
   
  
   paired_data = NULL
   
   for (tissue_type_selected in tissue_types) {
     
     paired_samples = tumor_normal_paired_fusion_counts %>% filter(tissue_type == tissue_type_selected) %>% spread(key=dataclassTN, value=fusion_count, fill=0)  
     
     
     normal_counts = paired_samples %>% pull(`TCGA-normal`)
     tumor_counts = paired_samples %>% pull(`TCGA-tumor`)
     
     try( {
       
       t = t.test(normal_counts, tumor_counts, paired = T, alternative = 'less')
       tumor_normal_compare_pvalues = bind_rows(tumor_normal_compare_pvalues,
                                                data.frame(tissue_type=tissue_type_selected, 
                                                           min_ffpm = min_ffpm_use, tp_value=t$p.value))
       
       paired_data = bind_rows(paired_data,
                               data.frame(tissue_type=tissue_type_selected, type='normal', counts=normal_counts),
                               data.frame(tissue_type=tissue_type_selected, type='tumor', counts=tumor_counts))
     }) 
   }
   
   p = paired_data %>%
     ggplot(aes(x=tissue_type, y=counts, fill=type)) + geom_split_violin(scale='width') + 
     theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(1))) +
     ggtitle(paste("TCGA fusion T/N paired counts at min FFPM threshold: ", min_ffpm_use))  + geom_jitter(size=rel(0.1))
   #+ ylim(0,50)
   
   plot(p)
   
}



```

Below we investigate the statistical significance of differences in the violin plots

```{r}

# correct p-values for mult tests
tumor_normal_compare_pvalues$tp_pval_BH = p.adjust(tumor_normal_compare_pvalues$tp_value, method="BH")

tumor_normal_compare_pvalues  = tumor_normal_compare_pvalues %>% 
  mutate(signif = tp_pval_BH < 0.05) %>%
  mutate(pt_negLogPval = -1 * log10(tp_pval_BH)) 

p = tumor_normal_compare_pvalues %>% 
  ggplot(aes(x=tissue_type, y=pt_negLogPval, fill=signif)) + facet_wrap(~min_ffpm) + 
  geom_col() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(1)))

plot(p)

```

Examine number of tumor samples having significant differences given the minimum FFPM 

```{r}

tumor_normal_compare_pvalues %>% filter(signif) %>% group_by(min_ffpm) %>% tally() %>%
  ggplot(aes(x=min_ffpm, y=n)) + geom_point() + geom_line()

```




## Examine distribution of fusion expression levels

Below we explore the distribution of fusion expression levels according to the classes of fusions, and examine the subset of COSMIC fusions in comparison

- used for Figure 4b

```{r}

data2 = data %>% mutate(dataclass2 = dataclassTN) %>% 
    mutate(dataclass2 = ifelse(grepl("Cosmic",Annots), "Cosmic", dataclass2))

cosmic_fusion_FFPM = data2 %>% filter(dataclass2 == "Cosmic") %>% pull(FFPM)
GTEx_normal_FFPM = data2 %>% filter(dataclassTN == "GTEx-normal") %>% pull(FFPM)
TCGA_tumor_FFPM = data2 %>% filter(dataclassTN == "TCGA-tumor") %>% pull(FFPM)
TCGA_normal_FFPM = data2 %>% filter(dataclassTN == "TCGA-normal") %>% pull(FFPM)

df_FFPM = bind_rows( data.frame(dataclass="Cosmic", FFPM=cosmic_fusion_FFPM),
                     data.frame(dataclass="GTEx-normal", FFPM=GTEx_normal_FFPM),
                     data.frame(dataclass="TCGA-tumor", FFPM=TCGA_tumor_FFPM),
                     data.frame(dataclass="TCGA-normal", FFPM=TCGA_normal_FFPM) ) 

                     
p = df_FFPM %>% mutate(FFPM_max1 = pmin(1, FFPM)) %>% ggplot(aes(x=FFPM_max1, color=dataclass)) + geom_density(alpha=0.2) + xlim(0, 1) + xlab("Density") + ylab("FFPM") +
  theme(legend.title = element_text(size = rel(0.7)),
         legend.text  = element_text(size = rel(0.7)),
         text=element_text(family="Times New Roman"))

plot(p)

ggsave("fuson_expr_density.svg", p, width=7.65, height=5.1)

```

Compare ECDFs for the above.

- used for Figure 4c
```{r}

cosmic_fusion_FFPM_ecdf = ecdf(cosmic_fusion_FFPM)
GTEx_normal_FFPM_ecdf = ecdf(GTEx_normal_FFPM)
TCGA_tumor_FFPM_ecdf = ecdf(TCGA_tumor_FFPM)
TCGA_normal_FFPM_ecdf = ecdf(TCGA_normal_FFPM)

ecdf_range = seq(0, 3, 0.01)

df_ecdf = bind_rows( data.frame(dataclass="Cosmic", minFFPM=ecdf_range, cum_frac=cosmic_fusion_FFPM_ecdf(ecdf_range)),
                     data.frame(dataclass="GTEx-normal", minFFPM=ecdf_range, cum_frac=GTEx_normal_FFPM_ecdf(ecdf_range)),
                     data.frame(dataclass="TCGA-tumor", minFFPM=ecdf_range, cum_frac=TCGA_tumor_FFPM_ecdf(ecdf_range)),
                     data.frame(dataclass="TCGA-normal", minFFPM=ecdf_range, cum_frac=TCGA_normal_FFPM_ecdf(ecdf_range)) )



p = df_ecdf %>% ggplot(aes(x=minFFPM, y=cum_frac, color=dataclass)) + geom_point() + 
       xlab("Minimum FFPM") + ylab("Cumulative Fraction") +
   theme(legend.title = element_text(size = rel(0.7)),
         legend.text  = element_text(size = rel(0.7)),
         text=element_text(family="Times New Roman"))
plot(p)

ggsave("fuson_cumul_expr_frac.svg", p, width=7.65, height=5.1)

```

```{r}
message("GTEx ecdf at 0.1 ffpm: ", GTEx_normal_FFPM_ecdf(0.1))

message("cosmic fusions ecdf at 0.1 ffpm: ", cosmic_fusion_FFPM_ecdf(0.1))

```

```{r}

num_gtex_ge = length(GTEx_normal_FFPM[GTEx_normal_FFPM >= 0.1])
num_gtex_lt = length(GTEx_normal_FFPM[GTEx_normal_FFPM < 0.1])

num_cosmic_ge = length(cosmic_fusion_FFPM[cosmic_fusion_FFPM >= 0.1])
num_cosmic_lt = length(cosmic_fusion_FFPM[cosmic_fusion_FFPM < 0.1])

m = matrix(c(num_gtex_ge, num_gtex_lt, num_cosmic_ge, num_cosmic_lt), nrow=2, byrow = T)
fisher.test(m)


```


## Examine the COSMIC fusions:

Below we count the fraction of tumor or normal samples that correspond to COSMIC fusions

- used for Figure 4a

```{r}

cosmic_fusion_occurrence = data2 %>% filter(data_class == "GTEx" | tumor_or_normal == "tumor") %>% 
    filter(dataclass2 == "Cosmic") %>% 
    select(fusion_name, sample_name, data_class, tissue_type) %>% 
    unique() %>% 
    mutate(sample_tissue = paste(data_class, tissue_type)) %>% 
    select (fusion_name, sample_tissue) %>% 
    group_by(fusion_name, sample_tissue) %>% 
    tally()


sample_tissue_counts = data2 %>% filter(data_class == "GTEx" | tumor_or_normal == "tumor") %>% 
    select(sample_name, data_class, tissue_type) %>% 
    mutate(sample_tissue = paste(data_class, tissue_type)) %>% 
    select(sample_name, sample_tissue) %>% 
    unique() %>% 
    group_by(sample_tissue) %>% 
    tally()

sample_tissue_counts = sample_tissue_counts %>% rename(sample_tissue_count = n)

cosmic_fusion_occurrence = left_join(x=cosmic_fusion_occurrence, y=sample_tissue_counts, by='sample_tissue')

cosmic_fusion_occurrence = cosmic_fusion_occurrence %>% mutate(sample_pct = n / sample_tissue_count * 100)

ranked_fusions = cosmic_fusion_occurrence %>% 
    group_by(fusion_name) %>% 
    summarize(sumpct= sum(sample_pct)) %>% 
    arrange(desc(sumpct)) %>% 
    select(fusion_name)
 

p = cosmic_fusion_occurrence %>% 
    ggplot(aes(x=factor(fusion_name, levels=ranked_fusions$fusion_name), y=sample_pct, fill=sample_tissue)) + 
    geom_col() + 
    theme(axis.text.x=element_text(angle = 90, hjust = 0, size=rel(0.75))) +
    xlab("COSMIC Fusion") + ylab("% of Samples") +
   theme(legend.title = element_text(size = rel(0.7)),
         legend.text  = element_text(size = rel(0.7)),
         text=element_text(family="Times New Roman"))

plot(p)

ggsave("cosmic_pct_samples.svg", p, width=7.65, height=5.1)
```


when we find cosmic fusions in normal GTEx, at what fractions of samples?

```{r}


cosmic_fusion_occurrence %>% filter(grepl("GTEx", sample_tissue)) %>% arrange(desc(sample_pct))
```



Examine distribution of FFPM for cosmic fusions

- used for Figure 4d

```{r}


p = data2 %>% filter(dataclass2 == "Cosmic") %>% 
    ggplot(aes(x=reorder(fusion_name, FFPM, FUN = median), y=FFPM)) + 
    geom_boxplot() +
     theme(axis.text.x=element_text(angle = 90, hjust = 0, size=rel(0.7))) + 
    scale_y_continuous(trans='log2') + 
    #ggtitle("Cosmic Fusion Expression Distribution") +
    xlab("COSMIC Fusion") +
    theme(legend.title = element_text(size = rel(0.7)),
         legend.text  = element_text(size = rel(0.7)),
         text=element_text(family="Times New Roman"))

plot(p)

ggsave("cosmic_ffpm_dist.svg", p, width=7.65, height=5.1)

```




