---
title: "STAR-Fusion_v1.7_TCGA_n_GTEx"
author: "Brian Haas"
date: "2/9/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(extrafont)
```


```{r}
data = read.table(gzfile("../data/TCGA_n_GTEx.STAR-Fusion.v1.7.fusion_preds.decorated.tsv.gz"), header=T, sep="\t", stringsAsFactors = F)

data = data %>% mutate(dataclassTN = paste0(data_class, "-", tumor_or_normal))

#Count numbers of samples before basic filtering

data %>% select(dataclassTN, sample_name) %>% unique() %>% ungroup() %>% group_by(dataclassTN) %>% tally()

## some initial cleanup
data = data %>% 
  filter(! grepl("chrM", Annots)) %>%
  filter(! grepl("HLA", fusion_name)) %>%
  filter(! (grepl("^IG[HKVL]", LeftGene) & grepl("^IG[HKVL]", RightGene)))

## only consider the primary isoform.
data = data %>% filter(is_primary)


# require reference splicing
data = data %>% filter(SpliceType == "ONLY_REF_SPLICE")


#Count numbers of samples post basic filtering

data %>% select(dataclassTN, sample_name) %>% unique() %>% ungroup() %>% group_by(dataclassTN) %>% tally()

```




```{r}

get_FPM_vs_min_FFPM_via_tissue_type = function (data, min_ffpm_vals = seq(0, 0.3, 0.025)) {
    
    fusion_counts_vs_min_ffpm = NULL
    
    
    # ignore READTHRU fusions here
    
    samples_n_tissues = data %>%   
      filter(structure_type != "READTHRU") %>% 
      select(data_class, dataclassTN, tissue_type, sample_name) %>% unique()
    
    
    for (min_ffpm in min_ffpm_vals) {
        message(min_ffpm)
        
        data_filt = data %>% filter(FFPM >= min_ffpm)
        
        counts_per_filt_sample = data_filt %>% select(data_class, dataclassTN, tissue_type, sample_name, fusion_name) %>% 
            unique() %>% group_by(data_class, dataclassTN, tissue_type, sample_name) %>% tally(name='fusion_count')
        
        counts_per_filt_sample = left_join(samples_n_tissues, counts_per_filt_sample, 
                                           by=c('data_class', 'dataclassTN', 'tissue_type', 'sample_name')) %>%
        mutate(fusion_count = ifelse(is.na(fusion_count), 0, fusion_count))
        
        
        
        counts_per_filt_sample$min_ffpm = min_ffpm
        
        fusion_counts_vs_min_ffpm = rbind(fusion_counts_vs_min_ffpm, counts_per_filt_sample)
        
    }
    
    return(fusion_counts_vs_min_ffpm)
}


```


```{r}



tissue_fusion_counts_vs_min_ffpm = get_FPM_vs_min_FFPM_via_tissue_type(data)


```


Compute median fusion count for each tissue type

```{r}
median_tissue_fusion_counts = tissue_fusion_counts_vs_min_ffpm  %>% 
    group_by(data_class, dataclassTN, tissue_type, min_ffpm) %>% 
    summarize(median_fusion_count = median(fusion_count)) %>% ungroup()

median_tissue_fusion_counts 
```

Counts of GTEx-normal fusions vs. min FFPM

```{r}
median_tissue_fusion_counts %>% filter(dataclassTN == "GTEx-normal") %>% 
    ggplot(aes(x=min_ffpm, y=median_fusion_count, color=tissue_type)) + geom_point() + geom_line()

```

Counts of TCGA-tumor fusions vs. min FFPM

```{r}
median_tissue_fusion_counts %>% filter(dataclassTN == "TCGA-tumor") %>% 
    ggplot(aes(x=min_ffpm, y=median_fusion_count, color=tissue_type)) + geom_point() + geom_line()

```


Jitter plots for counts of fusions per tissue type according to data class (ie. tcga-tumor, tcga-normal, and gtex-normal)
```{r}


min_ffpms = median_tissue_fusion_counts %>% select(min_ffpm) %>% unique() %>% arrange(min_ffpm) %>% pull(min_ffpm)

median_count_stats = NULL

for (min_ffpm_use in min_ffpms) {
    
    print(paste0("Min FFPM: ", min_ffpm_use))
    
    median_counts_min_ffpm  = median_tissue_fusion_counts %>% filter(min_ffpm == min_ffpm_use)
    
    p = median_counts_min_ffpm %>%
        ggplot(aes(x = dataclassTN, y=median_fusion_count)) + 
        geom_point(position = position_jitter(w = 0.1, h = 0)) +
        ggtitle(paste("median fusion counts/sample with min FFPM=", min_ffpm_use))
    
    
    GTEx_normal_counts = median_counts_min_ffpm %>% filter(dataclassTN == "GTEx-normal") %>% pull(median_fusion_count)
    
    TCGA_normal_counts = median_counts_min_ffpm %>% filter(dataclassTN == "TCGA-normal") %>% pull(median_fusion_count)
    
    TCGA_tumor_counts = median_counts_min_ffpm %>% filter(dataclassTN == "TCGA-tumor") %>% pull(median_fusion_count)
    
    print("GTEx-normal vs. TCGA-normal")
    t = t.test(GTEx_normal_counts, TCGA_normal_counts)
    print(t)
    median_count_stats = bind_rows(median_count_stats, data.frame(min_ffpm=min_ffpm_use, comparison = "GTEx-normal vs. TCGA-normal", tp_value = t$p.value))
    
    print("GTEx-normal vs. TCGA-tumor")
    t = t.test(GTEx_normal_counts, TCGA_tumor_counts)
    print(t)
    median_count_stats = bind_rows(median_count_stats, data.frame(min_ffpm=min_ffpm_use, comparison = "GTEx-normal vs. TCGA-tumor", tp_value = t$p.value ))
    
    print("TCGA-normal vs. TCGA-tumor")
    t = t.test(TCGA_normal_counts, TCGA_tumor_counts)
    print(t)
    median_count_stats = bind_rows(median_count_stats, data.frame(min_ffpm=min_ffpm_use, comparison = "TCGA-normal vs. TCGA-tumor", tp_value = t$p.value  ))
    
    plot(p)
    
}
```


```{r}

median_count_stats = median_count_stats %>% mutate(tnegLogP = -1 * log(tp_value))
median_count_stats %>% ggplot(aes(x=min_ffpm, y=tnegLogP, color=comparison)) + geom_point() + geom_line() + ggtitle("t-test -Log(p)")

median_count_stats 

```

```{r}

median_tissue_fusion_counts %>% filter(min_ffpm == 0.1) %>% group_by(dataclassTN) %>% summarize(median(median_fusion_count))

```

```{r}

median_tissue_fusion_counts %>% filter(min_ffpm == 0.1) %>% filter(tissue_type == "Pancreas")

```





Plot TCGA tumor/normal median fusion counts according to matched sample types.

```{r}

tcga_tissues_with_matched_normals = median_tissue_fusion_counts %>% filter(dataclassTN == "TCGA-normal") %>% 
    select(tissue_type) %>% unique() %>% pull(tissue_type)


for (min_ffpm_use in min_ffpms) {
    
  # statistical analysis of medians
  tumor_normal_paired_medians = median_tissue_fusion_counts %>% 
    filter(data_class == "TCGA" & tissue_type %in% tcga_tissues_with_matched_normals & min_ffpm == min_ffpm_use) %>%
    arrange(tissue_type) %>% spread(key=dataclassTN, value=median_fusion_count)
  
   normal_medians = tumor_normal_paired_medians %>% select(`TCGA-normal`) %>% pull(`TCGA-normal`)
   tumor_medians = tumor_normal_paired_medians %>% select(`TCGA-tumor`) %>% pull(`TCGA-tumor`)
   t = t.test(tumor_medians, normal_medians, paired = T)
   print(min_ffpm_use)
   print(t)
  
   
   p = median_tissue_fusion_counts %>% filter(data_class == "TCGA" & tissue_type %in% tcga_tissues_with_matched_normals) %>%
     filter(min_ffpm == min_ffpm_use) %>%
     ggplot(aes(x=tissue_type, y=median_fusion_count, fill=dataclassTN)) + geom_col(position='dodge') + 
     theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(1))) +
     ggtitle(paste("Median TCGA fusion counts at min FFPM threshold: ", min_ffpm_use))
   
   
   
       
   plot(p)
   
}

```



```{r}

source("../Rlib/geom_split_violin.R")

tcga_sample_info = data %>% filter(data_class == "TCGA") %>% select(tissue_type, sample_name) %>% unique()
tcga_sample_info$patient = str_replace(tcga_sample_info$sample_name, "-(NT|TP)$", "")
patients_with_both_tumor_normal_samples = tcga_sample_info %>% group_by(patient) %>% tally() %>% filter(n>1) %>% pull(patient)

tissue_fusion_counts_vs_min_ffpm$patient = str_replace(tissue_fusion_counts_vs_min_ffpm$sample_name, "-(NT|TP)$", "")

tumor_normal_compare_pvalues = NULL

for (min_ffpm_use in min_ffpms) {
    
  # statistical analysis of paired counts
  tumor_normal_paired_fusion_counts = tissue_fusion_counts_vs_min_ffpm %>% 
    filter(patient %in% patients_with_both_tumor_normal_samples) %>%
    filter(data_class == "TCGA" & tissue_type %in% tcga_tissues_with_matched_normals & min_ffpm == min_ffpm_use) %>%
    select(data_class, dataclassTN, patient, tissue_type, fusion_count) %>%
    arrange(tissue_type) 

  
   tissue_types = tumor_normal_paired_fusion_counts %>% select(tissue_type) %>% unique() %>% pull(tissue_type)
   
  
   paired_data = NULL
   
   for (tissue_type_selected in tissue_types) {
     
     paired_samples = tumor_normal_paired_fusion_counts %>% filter(tissue_type == tissue_type_selected) %>% spread(key=dataclassTN, value=fusion_count, fill=0)  
     
     
     normal_counts = paired_samples %>% pull(`TCGA-normal`)
     tumor_counts = paired_samples %>% pull(`TCGA-tumor`)
     
     try( {
       
       t = t.test(normal_counts, tumor_counts, paired = T, alternative = 'less')
       tumor_normal_compare_pvalues = bind_rows(tumor_normal_compare_pvalues,
                                                data.frame(tissue_type=tissue_type_selected, 
                                                           min_ffpm = min_ffpm_use, tp_value=t$p.value))
       
       paired_data = bind_rows(paired_data,
                               data.frame(tissue_type=tissue_type_selected, type='normal', counts=normal_counts),
                               data.frame(tissue_type=tissue_type_selected, type='tumor', counts=tumor_counts))
     }) 
   }
   
   p = paired_data %>%
     ggplot(aes(x=tissue_type, y=counts, fill=type)) + geom_split_violin(scale='width') + 
     theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(1))) +
     ggtitle(paste("TCGA fusion T/N paired counts at min FFPM threshold: ", min_ffpm_use))  + geom_jitter(size=rel(0.1))
   #+ ylim(0,50)
   
   plot(p)
   
}



```

```{r}

# correct p-values for mult tests
tumor_normal_compare_pvalues$tp_pval_BH = p.adjust(tumor_normal_compare_pvalues$tp_value, method="BH")

tumor_normal_compare_pvalues  = tumor_normal_compare_pvalues %>% 
  mutate(signif = tp_pval_BH < 0.05) %>%
  mutate(pt_negLogPval = -1 * log10(tp_pval_BH)) 

p = tumor_normal_compare_pvalues %>% 
  ggplot(aes(x=tissue_type, y=pt_negLogPval, fill=signif)) + facet_wrap(~min_ffpm) + 
  geom_col() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(1)))

plot(p)

```


```{r}

tumor_normal_compare_pvalues %>% filter(signif) %>% group_by(min_ffpm) %>% tally() %>%
  ggplot(aes(x=min_ffpm, y=n)) + geom_point() + geom_line()

```


```{r}


for (min_ffpm_use in min_ffpms) {

    p =  median_tissue_fusion_counts %>% 
        filter(min_ffpm == min_ffpm_use) %>% 
        group_by(dataclassTN, min_ffpm) %>% 
        summarize(median_of_medians = median(median_fusion_count)) %>%
        ggplot(aes(x=dataclassTN, y=median_of_medians)) + geom_col() + 
        ggtitle(paste("Median of median fusion counts, min ffpm = ", min_ffpm_use))
    
    plot(p)

}

```


# Examine distribution of fusion expression levels

```{r}

data2 = data %>% mutate(dataclass2 = dataclassTN) %>% 
    mutate(dataclass2 = ifelse(grepl("Cosmic",Annots), "Cosmic", dataclass2))

cosmic_fusion_FFPM = data2 %>% filter(dataclass2 == "Cosmic") %>% pull(FFPM)
GTEx_normal_FFPM = data2 %>% filter(dataclassTN == "GTEx-normal") %>% pull(FFPM)
TCGA_tumor_FFPM = data2 %>% filter(dataclassTN == "TCGA-tumor") %>% pull(FFPM)
TCGA_normal_FFPM = data2 %>% filter(dataclassTN == "TCGA-normal") %>% pull(FFPM)

df_FFPM = bind_rows( data.frame(dataclass="Cosmic", FFPM=cosmic_fusion_FFPM),
                     data.frame(dataclass="GTEx-normal", FFPM=GTEx_normal_FFPM),
                     data.frame(dataclass="TCGA-tumor", FFPM=TCGA_tumor_FFPM),
                     data.frame(dataclass="TCGA-normal", FFPM=TCGA_normal_FFPM) ) 

```

```{r}
                     
p = df_FFPM %>% mutate(FFPM_max1 = pmin(1, FFPM)) %>% ggplot(aes(x=FFPM_max1, color=dataclass)) + geom_density(alpha=0.2) + xlim(0, 1) + xlab("Density") + ylab("FFPM") +
  theme(legend.title = element_text(size = rel(0.7)),
         legend.text  = element_text(size = rel(0.7)),
         text=element_text(family="Times New Roman"))

plot(p)

ggsave("fuson_expr_density.svg", p, width=7.65, height=5.1)

```

Compare ECDFs for the above.
```{r}

cosmic_fusion_FFPM_ecdf = ecdf(cosmic_fusion_FFPM)
GTEx_normal_FFPM_ecdf = ecdf(GTEx_normal_FFPM)
TCGA_tumor_FFPM_ecdf = ecdf(TCGA_tumor_FFPM)
TCGA_normal_FFPM_ecdf = ecdf(TCGA_normal_FFPM)

ecdf_range = seq(0, 3, 0.01)

df_ecdf = bind_rows( data.frame(dataclass="Cosmic", minFFPM=ecdf_range, cum_frac=cosmic_fusion_FFPM_ecdf(ecdf_range)),
                     data.frame(dataclass="GTEx-normal", minFFPM=ecdf_range, cum_frac=GTEx_normal_FFPM_ecdf(ecdf_range)),
                     data.frame(dataclass="TCGA-tumor", minFFPM=ecdf_range, cum_frac=TCGA_tumor_FFPM_ecdf(ecdf_range)),
                     data.frame(dataclass="TCGA-normal", minFFPM=ecdf_range, cum_frac=TCGA_normal_FFPM_ecdf(ecdf_range)) )


```

```{r}
p = df_ecdf %>% ggplot(aes(x=minFFPM, y=cum_frac, color=dataclass)) + geom_point() + 
       xlab("Minimum FFPM") + ylab("Cumulative Fraction") +
   theme(legend.title = element_text(size = rel(0.7)),
         legend.text  = element_text(size = rel(0.7)),
         text=element_text(family="Times New Roman"))
plot(p)

ggsave("fuson_cumul_expr_frac.svg", p, width=7.65, height=5.1)

```

```{r}
message("GTEx ecdf at 0.1 ffpm: ", GTEx_normal_FFPM_ecdf(0.1))

message("cosmic fusions ecdf at 0.1 ffpm: ", cosmic_fusion_FFPM_ecdf(0.1))

```

```{r}

num_gtex_ge = length(GTEx_normal_FFPM[GTEx_normal_FFPM >= 0.1])
num_gtex_lt = length(GTEx_normal_FFPM[GTEx_normal_FFPM < 0.1])

num_cosmic_ge = length(cosmic_fusion_FFPM[cosmic_fusion_FFPM >= 0.1])
num_cosmic_lt = length(cosmic_fusion_FFPM[cosmic_fusion_FFPM < 0.1])

m = matrix(c(num_gtex_ge, num_gtex_lt, num_cosmic_ge, num_cosmic_lt), nrow=2, byrow = T)
fisher.test(m)


```


## Examine the COSMIC fusions:

```{r}

cosmic_fusion_occurrence = data2 %>% filter(data_class == "GTEx" | tumor_or_normal == "tumor") %>% 
    filter(dataclass2 == "Cosmic") %>% 
    select(fusion_name, sample_name, data_class, tissue_type) %>% 
    unique() %>% 
    mutate(sample_tissue = paste(data_class, tissue_type)) %>% 
    select (fusion_name, sample_tissue) %>% 
    group_by(fusion_name, sample_tissue) %>% 
    tally()


sample_tissue_counts = data2 %>% filter(data_class == "GTEx" | tumor_or_normal == "tumor") %>% 
    select(sample_name, data_class, tissue_type) %>% 
    mutate(sample_tissue = paste(data_class, tissue_type)) %>% 
    select(sample_name, sample_tissue) %>% 
    unique() %>% 
    group_by(sample_tissue) %>% 
    tally()

sample_tissue_counts = sample_tissue_counts %>% rename(sample_tissue_count = n)

cosmic_fusion_occurrence = left_join(x=cosmic_fusion_occurrence, y=sample_tissue_counts, by='sample_tissue')

cosmic_fusion_occurrence = cosmic_fusion_occurrence %>% mutate(sample_pct = n / sample_tissue_count * 100)

ranked_fusions = cosmic_fusion_occurrence %>% 
    group_by(fusion_name) %>% 
    summarize(sumpct= sum(sample_pct)) %>% 
    arrange(desc(sumpct)) %>% 
    select(fusion_name)
 
```

```{r}

p = cosmic_fusion_occurrence %>% 
    ggplot(aes(x=factor(fusion_name, levels=ranked_fusions$fusion_name), y=sample_pct, fill=sample_tissue)) + 
    geom_col() + 
    theme(axis.text.x=element_text(angle = 90, hjust = 0, size=rel(0.75))) +
    xlab("COSMIC Fusion") + ylab("% of Samples") +
   theme(legend.title = element_text(size = rel(0.7)),
         legend.text  = element_text(size = rel(0.7)),
         text=element_text(family="Times New Roman"))

plot(p)

ggsave("cosmic_pct_samples.svg", p, width=7.65, height=5.1)
```


when we find cosmic fusions in normal GTEx, at what fractions of samples?

```{r}


cosmic_fusion_occurrence %>% filter(grepl("GTEx", sample_tissue)) %>% arrange(desc(sample_pct))
```



Examine distribution of FFPM for cosmic fusions

```{r}


p = data2 %>% filter(dataclass2 == "Cosmic") %>% 
    ggplot(aes(x=reorder(fusion_name, FFPM, FUN = median), y=FFPM)) + 
    geom_boxplot() +
     theme(axis.text.x=element_text(angle = 90, hjust = 0, size=rel(0.7))) + 
    scale_y_continuous(trans='log2') + 
    #ggtitle("Cosmic Fusion Expression Distribution") +
    xlab("COSMIC Fusion") +
    theme(legend.title = element_text(size = rel(0.7)),
         legend.text  = element_text(size = rel(0.7)),
         text=element_text(family="Times New Roman"))

plot(p)

ggsave("cosmic_ffpm_dist.svg", p, width=7.65, height=5.1)

```




