---
title: "Scaling_and_Clustering_Fusions"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(umap)
library(leiden)
library(cowplot)
library(ggrepel)
library(FNN)
library(igraph)
library(pheatmap)
```


Below demonstrates the general process by which we scaled fusion attribute values, generated fusion clusters, and examined the content of those clusters.

Note, it isn't clear that this process can be run as fully reproducible. Running it on different machines has yielded different results - however, the numbers of clusters defined are similar and the enrichments we observe for COSMIC fusions are consistent.

- See ../3.Annotate_Leiden_Fusion_Clusters/AnnotateLeidenClusters.Rmd for how we analyzed our UMAP ordination and defined clusters.


```{r}

fusion_data = read.table(gzfile("../data/FusionInspector.v2.4.0.examine_recurrents.tsv.gz"), header=T, sep="\t", stringsAsFactors = F)

# remove the earlier umap and leiden assignments as we'll be recomputing them here.

fusion_data = fusion_data %>% select(-umap1, -umap2, -leiden)

```


```{r}
# prep data for scaling and clustering

pseudocount = 1
data = data.frame(adj_annot_splice=fusion_data$annot_splice, 
                  adj_consensus_splice=fusion_data$consensus_splice)

data$adj_FFPM=log2(fusion_data$FFPM+pseudocount)
data$adj_left_counter_ffpm = log2(fusion_data$left_counter_ffpm+pseudocount)
data$adj_right_counter_ffpm = log2(fusion_data$right_counter_ffpm+pseudocount)
data$adj_FAR_left = log2(fusion_data$FAR_left+pseudocount)
data$adj_FAR_right = log2(fusion_data$FAR_right+pseudocount)
data$adj_microh_brkpt_dist = log2(fusion_data$microh_brkpt_dist+1)
data$adj_num_microh = log2(fusion_data$num_microh+1)

```

```{r}
data %>% gather(key=key, value=value, 
                adj_annot_splice, adj_consensus_splice,
                adj_FFPM, adj_left_counter_ffpm, adj_right_counter_ffpm, 
                adj_FAR_left, adj_FAR_right, adj_microh_brkpt_dist, adj_num_microh) %>% ggplot(aes(x=key, y=value)) + geom_violin() +  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

```{r}
data.colmeans = colMeans(data)
message("data.colmeans")
data.colmeans

data.sd = apply(data, 2, sd)
message("data.sd")
data.sd


```


```{r}
data.scaled = scale(data, center=T, scale=T)
data.scaled.df = as.data.frame(data.scaled)
data.scaled.df %>% gather() %>% ggplot(aes(x=key, y=value)) + geom_violin() +  
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```



```{r}
# truncate to interval [-2,2]
data.scaled[data.scaled < -2] = -2
data.scaled[data.scaled > 2] = 2

data.scaled.df = as.data.frame(data.scaled)

as.data.frame(data.scaled) %>% gather() %>% ggplot(aes(x=key, y=value)) + geom_violin() +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```

Do another round of scaling to the -2,2 interval:


```{r}

apply(data.scaled, 2, range)

```


```{r}

rescale = function(vals, scale_range=c(-2,2)) {
  
  min_val = min(vals)
  max_val = max(vals)
  
  inrange = max_val - min_val
  outrange = scale_range[2] - scale_range[1]
  
  adj_vals = (vals - min_val) / inrange * outrange + scale_range[1]
  
  return(adj_vals)
}

data.scaled = apply(data.scaled, 2, rescale)

data.scaled.df = as.data.frame(data.scaled)

as.data.frame(data.scaled) %>% gather() %>% ggplot(aes(x=key, y=value)) + geom_violin() +
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))
```



## umap on data.scaled

```{r}

umap.mydefaults = umap.defaults
umap.mydefaults$random_state =  123456
data.scaled.umap = umap(data.scaled, config=umap.mydefaults)
#saveRDS(data.scaled.umap, file="data.scaled.umap.rds")
umap.data = data.scaled.umap$layout
colnames(umap.data) = c('umap1', 'umap2')
```



```{r}
as.data.frame(umap.data) %>% ggplot(aes(x = umap1, y = umap2)) + 
    geom_point(alpha = 0.3) + theme_bw() + ggtitle("umap data.scaled")
```


## Clustering Fusions 

```{r}

LEIDEN_RESOLUTION = 3.0

k = 50
knn = get.knn(data.scaled, k = k)

knn.dist = data.frame(
    from = rep(1:nrow(knn$nn.index), k), 
    to = as.vector(knn$nn.index), 
    weight = 1/(1 + as.vector(knn$nn.dist)))

knn.graph = graph_from_data_frame(knn.dist, directed = FALSE)
knn.graph.simple = simplify(knn.graph)


adjacency_matrix <- igraph::as_adjacency_matrix(knn.graph.simple)
partition <- leiden(adjacency_matrix, seed=12345, resolution_parameter = LEIDEN_RESOLUTION)
table(partition)


```

```{r}

# merge all back to the fusion_data table

fusion_data = cbind(fusion_data, data.scaled)
fusion_data = cbind(fusion_data, umap.data)
fusion_data$leiden = as.character(partition)

```


Plot umap for all fusion isoforms and highlight COSMIC isoforms

```{r}
umap.layout.cent = fusion_data %>% select(umap1, umap2, leiden) %>% group_by(leiden) %>% summarize_all(mean)


plot_umap = function() {

    p = fusion_data %>% ggplot(aes(x = umap1, y = umap2, colour = leiden)) + 
        geom_point(alpha = 0.3) + 
        theme_bw() + 
        geom_label_repel(data=umap.layout.cent, aes(label = leiden)) + 
                             guides(colour = FALSE)

    return(p)
}

umap_plot = plot_umap()

# highlight the cosmic entries

cosmic_subset = fusion_data %>% filter(cosmic == TRUE)

p = umap_plot + geom_point(data=cosmic_subset, color='black')

plot(p)


```

Restrict cosmic entries to just those highest expressed isoforms (primary isoforms) per sample.

```{r}
cosmic.primary_isoform = cosmic_subset %>% filter(is_primary)

p = umap_plot + geom_point(data=cosmic.primary_isoform, color='black')

plot(p)

```



Examine the median feature attribute for each cluster to understand differences between fusion clusters
```{r}

max_cluster = max(partition)

all_median_vals_df = NULL



for (cluster_i in seq(max_cluster)) {

    #message("cluster: ", cluster_i)
    
    cluster_row_indices = which(fusion_data$leiden == cluster_i)
    
    feature_atts = data.scaled[cluster_row_indices,]
    
    median_vals = as_tibble(feature_atts) %>% gather() %>% group_by(key) %>% summarize(m=median(value))

    df = data.frame(median_vals$m, row.names=median_vals$key)
    df = as.data.frame(t(df))
    df$cluster = cluster_i

    all_median_vals_df = bind_rows(all_median_vals_df, df)

}


rownames(all_median_vals_df) = all_median_vals_df$cluster

all_median_vals_df = all_median_vals_df[,!c(colnames(all_median_vals_df)=='cluster')]

median_vals_pca = prcomp(t(all_median_vals_df))
rows_d = dist(median_vals_pca$rotation[,1:6])
rowclust = hclust(as.dist(rows_d), "ward.D")


cols_d = dist(t(all_median_vals_df))

colclust = hclust(as.dist(cols_d), "ward.D")



p = pheatmap(all_median_vals_df, 
             cluster_rows=rowclust,
             cluster_cols=colclust,
             )

```


plot cosmic primary isoform fusion counts according to cluster

```{r}


cosmic_counts_per_cluster = cosmic.primary_isoform %>%
    group_by(fusion_name, leiden) %>% 
    tally()

cosmic_counts_per_cluster$leiden = factor(cosmic_counts_per_cluster$leiden, levels=rev(p$tree_row$labels[p$tree_row$order]))

empty_counts_df = data.frame(fusion_name="NA", leiden=factor(seq(max_cluster), levels=rev(p$tree_row$labels[p$tree_row$order])), n=0)

cosmic_counts_per_cluster = bind_rows(empty_counts_df, cosmic_counts_per_cluster)



# plot it

cosmic_sum_counts_per_cluster = cosmic_counts_per_cluster %>% group_by(leiden) %>% summarize(sum_n = sum(n))
cosmic_all_iso_counts_per_cluster_plot = cosmic_counts_per_cluster %>% ggplot(aes(x=leiden, y=n, fill=fusion_name)) + geom_col() + coord_flip() + ggtitle("COSMIC primary isoform count")



plot(cosmic_all_iso_counts_per_cluster_plot)
```


```{r}

# all fusions, top isoforms


top_fusion_counts_per_cluster = fusion_data %>% 
    filter(is_primary) %>%
    select(fusion_name, leiden) %>% 
    group_by(fusion_name, leiden) %>% 
    tally()

top_fusion_counts_per_cluster$leiden = factor(top_fusion_counts_per_cluster$leiden, levels=rev(p$tree_row$labels[p$tree_row$order]))

empty_counts_df = data.frame(fusion_name="NA", leiden=factor(seq(max_cluster), levels=rev(p$tree_row$labels[p$tree_row$order])), n=0)


top_fusion_counts_per_cluster = bind_rows(empty_counts_df, top_fusion_counts_per_cluster)

sum_fusion_counts_per_cluster = top_fusion_counts_per_cluster %>% group_by(leiden) %>% summarize(sum_n = sum(n))

sum_fusion_counts_per_cluster_plot = sum_fusion_counts_per_cluster %>% ggplot(aes(x=leiden, y=sum_n)) + geom_col() + coord_flip() + ggtitle("Fusion primary isoform count")

plot(sum_fusion_counts_per_cluster_plot)
```

plot cosmic enrichment per cluster

```{r}

cosmic_sum_counts_per_cluster$fusion_type = "cosmic"
sum_fusion_counts_per_cluster$fusion_type = "all"
cosmic_and_all_counts_df = bind_rows(cosmic_sum_counts_per_cluster, sum_fusion_counts_per_cluster)

cosmic_and_all_counts_df = cosmic_and_all_counts_df %>% spread(key=fusion_type, value=sum_n)
cosmic_and_all_counts_df = cosmic_and_all_counts_df %>% mutate(frac_cosmic = cosmic/all)

cosmic_frac_enrichment_plot = cosmic_and_all_counts_df %>% ggplot(aes(x=leiden, y=frac_cosmic)) + 
    geom_col() + coord_flip() + ggtitle("Fraction COSMIC fusions")
plot(cosmic_frac_enrichment_plot)
```



Create single summary plot

```{r}


cosmic_all_iso_counts_per_cluster_plot_NLeg = cosmic_all_iso_counts_per_cluster_plot + 
    theme(legend.position = "none")

pg = plot_grid(p[[4]], cosmic_all_iso_counts_per_cluster_plot_NLeg, sum_fusion_counts_per_cluster_plot, cosmic_frac_enrichment_plot, 
          nrow=1, rel_widths = c(3,1,1,1) )

plot(pg)

```






