---
title: "Examine_Clustered_TARGETdb_Fusions"
author: "bhaas"
date: '2022-09-04'
output: html_document
---

# Exploring all recurrent fusions in TARGET


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggrepel)
library(pheatmap)
library(cowplot)
library(ggforce)

```


```{r}

data = read.table(gzfile("../../data/TARGET_STAR-Fusion-v1.10.0_FusionInspector-v2.4.0.fusions.tsv.gz"), header=T, sep="\t", stringsAsFactors = F)



```



```{r}

# number distinct fusion gene pairs
data %>% select(fusion_name) %>% unique() %>% nrow()


# number distinct validated by FI
data %>% filter(FI_validated) %>% select(fusion_name) %>% unique() %>% nrow()


```


```{r}

# set top expressed fusion occurrence per participant as  primary isoform, preferentially choosing FI over STAR-F for same fusion variant

data = data %>% select(-sample) %>%
  group_by(participant, tumor_or_normal, fusion_name) %>% 
    arrange(desc(FFPM.FI), desc(FFPM.starF)) %>% 
      mutate(is_primary = row_number() == 1) %>% 
  ungroup()

message('all fusions')
data %>% group_by(FI_validated) %>% tally()

message('primary fusions')
data %>% filter(is_primary) %>% group_by(FI_validated) %>% tally()

```






```{r}

# are there any STAR-F COSMIC predictions not called by FI ?

data %>% filter(is_primary & ! FI_validated) %>% filter(cosmic)

# very few and each with very low FFPM
```



#####################################################################################################################################


############################################
# Explore Recurrent Fusions in TARGET
############################################

- exclude those found in normals within TARGET normal set
- exclude those found at least a few times in GTEx normals and are not known COSMIC


```{r}

fusions_found_in_normals = data %>% filter(tumor_or_normal == "normal") %>% select(fusion_name) %>% unique() %>% pull(fusion_name)

write.table(fusions_found_in_normals, "fusions_found_in_normals_STARF.tsv", sep="\t", row.names=F, col.names = F)

length(fusions_found_in_normals)
# 1229
```





```{r}

# get total fusion abundance across TARGET samples

primary_data = data %>% filter(is_primary) %>% select(-sample_type)

fusion_all_sample_counts = primary_data %>% group_by(fusion_name) %>% tally(name='tot_sample_count_w_fusion')

primary_data = left_join(primary_data, fusion_all_sample_counts, by='fusion_name')


primary_data %>% select(fusion_name, participant, tot_sample_count_w_fusion)
```





```{r}

# combine with earlier tcga/gtex occurrence stats
tcga_gtex_fusion_stats = read.table(gzfile("../../data/TCGA_n_GTEx.STAR-Fusion.v1.7.fusion_preds.stats.tsv.gz"), header=T, sep="\t", stringsAsFactors = F)

```


```{r}
primary_data  = left_join(primary_data, tcga_gtex_fusion_stats, by='fusion_name')
 
primary_data %>% select(fusion_name, participant, tot_sample_count_w_fusion, normal, tumor, logTN)
```



```{r}
# define the reciprocal fusions among these
# - the fusion_name A::B and B::A must show up in the same sample

reciprocal_fusions_data = primary_data %>% 
       group_by(participant, fusion_name, sorted_fusion_name) %>% unique() %>% group_by(participant, sorted_fusion_name) %>% tally() %>% filter(n>1) %>% ungroup()

# capture both A::B and B::A occurrences of fusions for these samples.
reciprocal_fusions_data = left_join(reciprocal_fusions_data, 
                                    primary_data, 
                                    by=c('participant', 'sorted_fusion_name'))



# examine fusion sample correlation matrix for only the reciprocal fusion subset

write.table(reciprocal_fusions_data %>% select(project, fusion_name) %>% unique(), 
  file="reciprocal.info.tsv", quote=F, sep="\t", col.names=F, row.names=F)



write.table(reciprocal_fusions_data %>% select(fusion_name, participant) %>%  mutate(found=1) %>% unique() %>% spread(key=participant, value=found, fill=0),
            file="reciprocal.sample_matrix.tsv", quote=F, sep="\t", row.names=F)


fusion_sample_matrix = read.table("reciprocal.sample_matrix.tsv", row.names=1, header=T, sep="\t")
write.table(t(fusion_sample_matrix), file="reciprocal.sample_matrix.tsv", quote=F, sep="\t", row.names=T)


# Make the sample correlation plot like so using PtR:
# trinityrnaseq/Analysis/DifferentialExpression/PtR -m reciprocal.sample_matrix.tsv -s reciprocal.info.tsv --heatmap --sample_cor_matrix --img_width 100 --img_height 100


# will use reciprocal fusion data later.

```


```{r}

# examine the cosmic-here set wrt our earlier tcga/gtex findings

primary_data %>% filter(cosmic) %>% select(fusion_name, tot_sample_count_w_fusion, normal, tumor, logTN) %>% unique()


```




```{r}

# DUX4 fusions

primary_data %>% filter(grepl("DUX4", fusion_name))


```


```{r}

# IGH fusions

primary_data %>% filter(grepl("IGH@-ext", fusion_name)) %>% group_by(fusion_name, tumor_or_normal) %>% 
  tally(name='count') %>% spread(key=tumor_or_normal, value=count) %>% arrange(desc(tumor))


```


```{r}

# CRLF2 fusions

primary_data %>% filter(grepl("CRLF2", fusion_name)) %>% group_by(fusion_name, tumor_or_normal) %>% tally(name='count') %>% spread(key=tumor_or_normal, value=count) %>% arrange(desc(tumor))


```


```{r}


primary_data %>% filter(fusion_name %in% c('P2RY8::CRLF2', 'IGH-@-ext::CRLF2') ) %>% arrange(desc(FFPM.starF))

```



#####################################
# select top TARGET recurrent fusions





```{r}

sample_type_counts = primary_data %>% select(participant, project, tumor_or_normal) %>% unique() %>% group_by(project, tumor_or_normal) %>% tally(name='sample_type_counts')

sample_type_counts 
```



```{r}


all_cosmic_fusions = primary_data %>% filter(cosmic) %>% select(fusion_name) %>% unique() %>% pull(fusion_name)

fusion_sample_type_counts = primary_data %>% 
  group_by(fusion_name, project, tumor_or_normal) %>% 
  tally(name='fusion_sample_type_counts')

fusion_sample_type_counts = left_join(fusion_sample_type_counts, sample_type_counts, by=c('project', 'tumor_or_normal'))


fusion_sample_type_counts = fusion_sample_type_counts %>% mutate(sample_type_frac = fusion_sample_type_counts / sample_type_counts)

fusion_sample_type_counts = fusion_sample_type_counts %>% mutate(cosmic = (fusion_name %in% all_cosmic_fusions))


fusion_sample_type_counts

```


```{r}
# examine the cosmic fractions


fusion_sample_type_counts %>% filter(cosmic) %>% arrange(desc(sample_type_frac))


```

```{r}
fusion_sample_type_counts %>% ungroup()  %>%  filter(cosmic) %>% select(fusion_name) %>% unique() %>% nrow()

```


```{r}

# find some new candidates of interest

top_new_candidates = primary_data %>%  
  filter(! fusion_name %in% fusions_found_in_normals) %>%
  filter(tot_sample_count_w_fusion >= 3) %>%
  filter(normal == 0 | is.na(normal) | cosmic) %>%
  arrange(desc(tot_sample_count_w_fusion))


top_new_candidates  %>% select(fusion_name, tot_sample_count_w_fusion) %>% unique() %>% arrange(desc(tot_sample_count_w_fusion))
```


```{r}

# counts of recurrent TARGET fusions

nrow(top_new_candidates) 
# 44555

top_new_candidates %>% select(fusion_name) %>% unique() %>% nrow()
# 5222


```



```{r}

# incorporate the sample type fractions

top_new_candidates = left_join(top_new_candidates, fusion_sample_type_counts,  by=c('fusion_name', 'cosmic', 'project', 'tumor_or_normal' ) )

top_new_candidates  %>%  select(-c(sorted_fusion_name, annots))
```




```{r}
sum_frac = top_new_candidates %>% 
  select(fusion_name, tot_sample_count_w_fusion, cosmic, sample_type_frac, project) %>% 
  unique() %>%
  group_by(fusion_name) %>% 
  mutate(sum_frac = sum(sample_type_frac)) %>% 
  select(fusion_name, sum_frac) %>% 
  unique()

top_new_candidates = left_join(top_new_candidates, sum_frac, by='fusion_name')
 
```

```{r}

# ensure at least one instance of the fusion was validated by FusionInspector
FI_validated_fusions = data %>% filter(FI_validated) %>% select(fusion_name) %>% unique() %>% pull(fusion_name)

top_new_candidates =  top_new_candidates %>% filter(fusion_name %in% FI_validated_fusions)

```



```{r}

sample_frac_ranked_fusions = top_new_candidates %>% select(project, fusion_name, sample_type_frac) %>% unique() %>% group_by(project) %>% arrange(desc(sample_type_frac)) %>% mutate(ranking=row_number()) %>% ungroup() 

sample_frac_ranked_fusions  %>%
  ggplot(aes(x=ranking, y=sample_type_frac)) + 
  geom_point(aes(color=project)) + geom_line(aes(color=project)) +
  scale_x_log10() +
  theme_bw() +
  geom_hline(yintercept = 0.10, color='black', linetype = 'dotted')  +
  geom_point(data = sample_frac_ranked_fusions %>% filter(fusion_name %in% all_cosmic_fusions), aes(x=ranking, y=sample_type_frac), color='black') +
     theme(axis.text.y = element_text(size=rel(0.7)))  +
        scale_fill_continuous(high = "#132B43", low = "#56B1F7", na.value="white") +
         theme(panel.border = element_blank(), panel.grid.major = element_blank(),
               panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))


```

```{r}

# how many recurrent (min 3) fusions do we have for each project?

sample_frac_ranked_fusions %>% group_by(project) %>% arrange(desc(ranking)) %>% filter(row_number() == 1)

```


```{r}
write.table(top_new_candidates, file="TARGET.all_candidates.tsv", quote=F, sep="\t", row.names = F)


write.table(top_new_candidates %>% select(fusion_name, project, fusion_sample_type_counts, sample_type_counts, sample_type_frac) %>% unique() %>% arrange(desc(sample_type_frac)), file="TARGET.all_candidates.summary.tsv", quote=F, sep="\t", row.names = F)

```



## Reduce top candidates according to min (sample_type_frac)

```{r}

MIN_SAMPLE_FRAC = 0.1

top_new_candidates %>%  filter(sample_type_frac >= MIN_SAMPLE_FRAC) %>% select(fusion_name) %>% unique()
  
```



```{r}

top_new_candidates_min_sample_frac = left_join(top_new_candidates %>% filter(sample_type_frac >= MIN_SAMPLE_FRAC) %>% select(fusion_name) %>% unique(),
                                               top_new_candidates,  
                                               by='fusion_name')


top_new_candidates_min_sample_frac %>% select(fusion_name, participant) %>% unique() %>% nrow()
```



```{r}

candidates_sample_frac_plot = top_new_candidates_min_sample_frac  %>% 
  select(fusion_name, project, sample_type_frac, sum_frac) %>% unique() %>%
  #group_by(fusion_name, project) %>% arrange(desc(sample_type_frac) %>% filter(row_number()==1) %>% ungroup() %>%
  ggplot(aes(x=reorder(fusion_name, -sum_frac), y=sample_type_frac, fill=project)) + geom_col() +  
  theme(axis.text.x = element_text(angle = 90, hjust = 1))

candidates_sample_frac_plot 

```


```{r}

# indicate if in cosmic or not

cosmic_entries_plot = top_new_candidates_min_sample_frac  %>% select(fusion_name, sum_frac, cosmic) %>% unique() %>%
  mutate(cosmic = ifelse(cosmic, TRUE, NA)) %>% 
  ggplot(aes(x=reorder(fusion_name, -sum_frac), y=TRUE, fill=cosmic)) + 
  geom_tile() +
  theme_bw() +
            theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank()) +
            theme(axis.title.y=element_blank(), axis.text.y = element_blank(), axis.ticks.y=element_blank()) +
            scale_fill_manual(na.value="white", values="purple") +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
               panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))

cosmic_entries_plot 



```


```{r}

# indicate if includes an ALL cancer gene or not

ALL_drivers_plot = top_new_candidates_min_sample_frac  %>% select(fusion_name, sum_frac, contains_ALL_driver_gene) %>% unique() %>%
  mutate(contains_ALL_driver_gene = ifelse(contains_ALL_driver_gene, TRUE, NA)) %>% 
  ggplot(aes(x=reorder(fusion_name, -sum_frac), y=TRUE, fill=contains_ALL_driver_gene)) + 
  geom_tile() +
  theme_bw() +
            theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank()) +
            theme(axis.title.y=element_blank(), axis.text.y = element_blank(), axis.ticks.y=element_blank()) +
            scale_fill_manual(na.value="white", values="cyan") +
  theme(panel.border = element_blank(), panel.grid.major = element_blank(),
               panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))

ALL_drivers_plot 



```







```{r}

# examine reciprocal fusions

# now further restrict to those fusions among our top candidates as earlier defined.
top_new_candidate_sorted_fusions = top_new_candidates_min_sample_frac  %>% select(sorted_fusion_name) %>% unique() %>% pull(sorted_fusion_name)

reciprocal_fusions_data = reciprocal_fusions_data %>% filter(sorted_fusion_name %in% top_new_candidate_sorted_fusions)

reciprocal_fusions_data %>% arrange(participant, sorted_fusion_name)




```

```{r}


top_new_candidates_min_sample_frac  = left_join(top_new_candidates_min_sample_frac , 
                               reciprocal_fusions_data %>%  select(fusion_name) %>% 
                                 unique() %>% mutate(reciprocal_fusion = TRUE),
                               by='fusion_name')

top_new_candidates_min_sample_frac  %>% filter(reciprocal_fusion) %>% arrange(sorted_fusion_name, fusion_name)

```

```{r}

reciprocal_fusion_plot = top_new_candidates_min_sample_frac  %>% select(fusion_name, sum_frac, reciprocal_fusion) %>% unique() %>%
  ggplot(aes(x=reorder(fusion_name, -sum_frac), y=TRUE, fill=reciprocal_fusion)) + 
  geom_tile() +
  theme_bw() +
            theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank()) +
            theme(axis.title.y=element_blank(), axis.text.y = element_blank(), axis.ticks.y=element_blank()) +
            scale_fill_manual(na.value="white", values="blue")  +
   theme(panel.border = element_blank(), panel.grid.major = element_blank(),
               panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))

reciprocal_fusion_plot

```







# Want COSMIC-like cluster stats and frame change stats too.



```{r}

top_new_candidate_fusions_min_sample_frac = top_new_candidates_min_sample_frac %>% select(fusion_name) %>% unique() %>% pull(fusion_name)


cosmic_like_pred_data = primary_data %>% filter(fusion_name %in% top_new_candidate_fusions_min_sample_frac) %>% 
  group_by(fusion_name, fusion_cluster_att) %>% tally() %>% mutate(frac_fusions=prop.table(n))

cosmic_like_pred_data %>% group_by(fusion_cluster_att) %>% tally()


cosmic_like_pred_data

```


```{r}


frac_leiden_C21 = primary_data %>% group_by(fusion_name, leiden) %>% tally() %>% mutate(frac_leiden_C21 = prop.table(n)) %>% ungroup() %>% filter(leiden==21)



```



```{r}


cosmic_like_pred_data = bind_rows(cosmic_like_pred_data %>% select(fusion_name, fusion_cluster_att, frac_fusions),
                                  
            frac_leiden_C21 %>% select(fusion_name, frac_fusions = frac_leiden_C21) %>% 
            unique() %>% 
            filter(fusion_name %in% top_new_candidate_fusions_min_sample_frac) %>%
            mutate(fusion_cluster_att = "frac_in_C21") )

cosmic_like_pred_data 

```



```{r}

ARTIFACT_CLUSTERS = c(71,79,78,73,77)

fusion_fraction_artifact_clusters = primary_data %>%  
  mutate(in_artifact_cluster = (leiden %in% ARTIFACT_CLUSTERS) ) %>%

  group_by(fusion_name, in_artifact_cluster ) %>% 
  tally(name='count_in_artcluster') %>% 
  mutate(frac_artcluster=prop.table(count_in_artcluster)) %>% 
  ungroup() %>% 
  filter(in_artifact_cluster)

fusion_fraction_artifact_clusters


```



```{r}

cosmic_like_pred_data = bind_rows(cosmic_like_pred_data,
                                  fusion_fraction_artifact_clusters %>% 
                                    filter(fusion_name %in%  top_new_candidates_min_sample_frac$fusion_name) %>%
                                    select(fusion_name, frac_fusions = frac_artcluster) %>% mutate(fusion_cluster_att = "artifactCluster" ) ) 
                                    
                                
```





```{r}


cosmic_like_pred_data$fusion_cluster_att = factor(cosmic_like_pred_data$fusion_cluster_att, levels=rev(c('Other', 'Untested', 'Artifact', "artifactCluster", 'COSMIC-like', 'frac_in_C21')))


cosmic_like_pred_plot = left_join(cosmic_like_pred_data, sum_frac) %>% 
  ggplot(aes(x=reorder(fusion_name, -sum_frac), y=fusion_cluster_att, fill=frac_fusions)) +
        geom_tile() +
        theme_bw() +
        theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank()) +
        ylab("Cluster Type") +
        theme(axis.text.y = element_text(size=rel(0.7)))  +
        scale_fill_continuous(high = "#132B43", low = "#56B1F7", na.value="white") +
         theme(panel.border = element_blank(), panel.grid.major = element_blank(),
               panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))

cosmic_like_pred_plot

```


The artifact prediction (from the RandomForest) and the actual artifactCluster (physically clustering in artifact-like enriched clusters) look identical.  Just keep  one for the plot.


```{r}

cosmic_like_pred_plot = left_join(cosmic_like_pred_data, sum_frac) %>% 
  filter(fusion_cluster_att != "artifactCluster") %>%
  ggplot(aes(x=reorder(fusion_name, -sum_frac), y=fusion_cluster_att, fill=frac_fusions)) +
        geom_tile() +
        theme_bw() +
        theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank()) +
        ylab("Cluster Type") +
        theme(axis.text.y = element_text(size=rel(0.7)))  +
        scale_fill_continuous(high = "#132B43", low = "#56B1F7", na.value="white") +
         theme(panel.border = element_blank(), panel.grid.major = element_blank(),
               panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))

cosmic_like_pred_plot


```



```{r}

fusion_coding_effects = read.table(gzfile("../../data/fusions.coding_effects.abridged.tsv.gz"), header=T, sep="\t", stringsAsFactors = F)


```


```{r}

candidate_fusion_coding_data = left_join(data %>% filter(is_primary) %>% filter(fusion_name %in% top_new_candidate_fusions_min_sample_frac),
                                         fusion_coding_effects,
                                         by=c('fusion_name', 'LeftBreakpoint', 'RightBreakpoint'))

candidate_fusion_coding_data 
```

```{r}

candidate_fusion_coding_frac_data = candidate_fusion_coding_data %>% select(fusion_name, PROT_FUSION_TYPE, participant) %>% 
  mutate(PROT_FUSION_TYPE = ifelse(PROT_FUSION_TYPE==".", "Other", PROT_FUSION_TYPE)) %>% 
  group_by(fusion_name, PROT_FUSION_TYPE) %>% 
  tally() %>% 
  mutate(frac_fusions=prop.table(n)) 

candidate_fusion_coding_frac_plot = left_join(candidate_fusion_coding_frac_data, sum_frac) %>%  
  ggplot(aes(x=reorder(fusion_name, -sum_frac), y=PROT_FUSION_TYPE, fill=frac_fusions)) +
        geom_tile() +
        theme_bw() +
        theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank()) +
        ylab("Cluster Type") +
        theme(axis.text.y = element_text(size=rel(0.7)))  +
        scale_fill_continuous(high = "#132B43", low = "#56B1F7", na.value="white")  +
         theme(panel.border = element_blank(), panel.grid.major = element_blank(),
               panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))


candidate_fusion_coding_frac_plot
```




```{r}
# chromosome structure


chrom_structure_plot =  top_new_candidates_min_sample_frac %>% 
  select(fusion_name, structure_type, sum_frac)  %>% 
  unique() %>%
    ggplot(aes(x=reorder(fusion_name, -sum_frac), y=0, fill=structure_type)) +
  geom_tile() +
   theme_bw() +
       theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank(),
          axis.title.y=element_blank(), axis.text.y=element_blank(), axis.ticks.y=element_blank()
          ) +
 
         theme(panel.border = element_blank(), panel.grid.major = element_blank(),
               panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))

chrom_structure_plot

```




```{r}
# combine plots into a single figure


pg = plot_grid(cosmic_like_pred_plot,
               candidate_fusion_coding_frac_plot,
               chrom_structure_plot,
               cosmic_entries_plot,
               reciprocal_fusion_plot,
               candidates_sample_frac_plot,
                 ncol=1,
                 align='v',
                 axis='lr',
                 rel_heights = c(0.2, 0.2, 0.15, 0.15, 0.15,0.70)
                 )


plot(pg)


```

```{r}

ggsave(pg, file="TARGET_summary_plot.pdf", height=8, width=14)


```

```{r}

# Make the short list:

top_new_candidates_top_sample_frac = top_new_candidates_min_sample_frac %>% 
  group_by(fusion_name) %>% 
  arrange(desc(sample_type_frac)) %>%
  filter(row_number()==1) %>%
  ungroup() %>%
  select(fusion_name, tot_sample_count_w_fusion, sample_type_frac, sum_frac, cosmic, normal, tumor, logTN, project, annots, structure_type, sorted_fusion_name) %>%
  arrange(desc(sample_type_frac))





top_new_candidates_top_sample_frac %>% arrange(desc(sample_type_frac))

```

```{r}
top_new_candidates_top_sample_frac %>% filter(cosmic)

```


```{r}
write.table(top_new_candidates_top_sample_frac, file=paste0("TARGET.recurrent_fusions_min", (MIN_SAMPLE_FRAC*100), "pct.tsv"), quote=F, sep="\t", row.names = F)

```



```{r}


blood_each_sample_type_heatmap = top_new_candidates_min_sample_frac %>% 

 ggplot(aes(x=reorder(fusion_name, -sum_frac), y=project)) +
        geom_tile(aes(fill=sample_type_frac)) +
        facet_col(vars(project), , scales = "free_y", space = "free") +

         theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank()) +
        ylab("sample_type") +
        xlab("fusion") +
        theme_bw() +
        theme(axis.text.y = element_text(size=rel(1)))  +
        scale_fill_continuous(high = "#132B43", low = "#56B1F7", na.value="white") +
        scale_x_discrete(expand = c(0, 0)) +    
        scale_y_discrete(expand = c(0, 0)) +  
        theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
        theme(panel.border = element_blank(), panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(), axis.line = element_line(colour = "black"))

plot(blood_each_sample_type_heatmap)

```

```{r}
ggsave(blood_each_sample_type_heatmap, file="blood_each_sample_type_heatmap.pdf", width=15, height=8)

```


```{r}
blood_top_sample_type_heatmap = top_new_candidates_min_sample_frac %>%

 ggplot(aes(x=reorder(fusion_name, -sum_frac), y=reorder(project, desc(project)))) +
        geom_tile(aes(fill=sample_type_frac)) +

#         theme(axis.title.x=element_blank(), axis.text.x=element_blank(), axis.ticks.x=element_blank())
#+
        ylab("sample_type") +
        xlab("fusion") +
        theme_bw() +
#        theme(axis.text.y = element_text(size=rel(1)))  +
        scale_fill_continuous(high = "#132B43", low = "#56B1F7", na.value="white") +
       theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
       theme(panel.border = element_blank(), 
             panel.grid.major = element_blank(),
               panel.grid.minor = element_blank(), 
             axis.line = element_line(colour = "black"))


plot(blood_top_sample_type_heatmap )
```




```{r}


pg = plot_grid(cosmic_like_pred_plot,
               candidate_fusion_coding_frac_plot,
               chrom_structure_plot,
               cosmic_entries_plot,
               ALL_drivers_plot, 
               reciprocal_fusion_plot,
               blood_top_sample_type_heatmap ,
                 ncol=1,
                 align='v',
                 axis='lr',
                 rel_heights = c(0.2, 0.2, 0.15, 0.1, 0.1, 0.1, 0.70)
                 )


plot(pg)

ggsave(pg, file=paste0("TARGET.recurrent_fusions_min", (MIN_SAMPLE_FRAC*100), "pct.pdf"), width=24, height=12)




```







