---
title: "AnnotateLeidenClusters"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(cowplot)
library(ggrepel)
```

Below we examine the contents of fusion clusters including COSMIC enrichment and properties of those COSMIC fusions found in COSMIC-enriched cluster C4.

- used for Figure 4


```{r}

fusion_data = read.table(gzfile("../data/FusionInspector.v2.4.0.examine_recurrents.tsv.gz"), header=T, sep="\t", stringsAsFactors = F)

fusion_data$leiden = factor(fusion_data$leiden)

umap.layout.cent = fusion_data %>% select(umap1, umap2, leiden) %>% group_by(leiden) %>% summarize_all(mean)


plot_umap = function() {

    p = fusion_data %>% ggplot(aes(x = umap1, y = umap2, colour = leiden)) + 
        geom_point(alpha = 0.3, size=rel(0.5)) + 
        theme_bw() + 
        geom_label_repel(data=umap.layout.cent, aes(label = leiden)) + 
                             guides(colour = FALSE)

    return(p)
}

umap_plot = plot_umap()

# highlight the cosmic entries

cosmic_subset = fusion_data %>% filter(cosmic & is_primary)

p = umap_plot + geom_point(data=cosmic_subset, color='black', size=rel(0.5))

plot(p)

ggsave("umap_res3.svg", p, width=9, height=6)

```

Generate heatmap of median cluster attributes

```{r}
# get median of scaled cluster attributes for heatmap

# prep data for scaling and clustering

pseudocount = 1
data = data.frame(adj_annot_splice=fusion_data$annot_splice, 
                  adj_consensus_splice=fusion_data$consensus_splice)

data$adj_FFPM=log2(fusion_data$FFPM+pseudocount)
data$adj_left_counter_ffpm = log2(fusion_data$left_counter_ffpm+pseudocount)
data$adj_right_counter_ffpm = log2(fusion_data$right_counter_ffpm+pseudocount)
data$adj_FAR_left = log2(fusion_data$FAR_left+pseudocount)
data$adj_FAR_right = log2(fusion_data$FAR_right+pseudocount)
data$adj_microh_brkpt_dist = log2(fusion_data$microh_brkpt_dist+1)
data$adj_num_microh = log2(fusion_data$num_microh+1)

# scale and center
data.scaled = scale(data, center=T, scale=T)

# truncate to interval [-2,2]
data.scaled[data.scaled < -2] = -2
data.scaled[data.scaled > 2] = 2


# Do another round of scaling to the -2,2 interval:

rescale = function(vals, scale_range=c(-2,2)) {
  
  min_val = min(vals)
  max_val = max(vals)
  
  inrange = max_val - min_val
  outrange = scale_range[2] - scale_range[1]
  
  adj_vals = (vals - min_val) / inrange * outrange + scale_range[1]
  
  return(adj_vals)
}

data.scaled = apply(data.scaled, 2, rescale)
data.scaled.df = as.data.frame(data.scaled)

as.data.frame(data.scaled) %>% gather() %>% ggplot(aes(x=key, y=value)) + geom_violin() +
   theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))


# get median of attribute values per cluster
clusters = fusion_data %>% select(leiden) %>% unique() %>% pull(leiden)
all_median_vals_df = NULL
for (cluster in clusters) {
    
    cluster_row_indices = which(fusion_data$leiden == cluster)
    
    feature_atts = data.scaled[cluster_row_indices,]
    
    median_vals = as_tibble(feature_atts) %>% gather() %>% group_by(key) %>% summarize(m=median(value))

    df = data.frame(median_vals$m, row.names=median_vals$key)
    df = as.data.frame(t(df))
    df$cluster = cluster

    all_median_vals_df = bind_rows(all_median_vals_df, df)

}

rownames(all_median_vals_df) = all_median_vals_df$cluster
all_median_vals_df = all_median_vals_df[,!c(colnames(all_median_vals_df)=='cluster')]


# generate heatmap, hierarchically clusters according to eigenvalues
median_vals_pca = prcomp(t(all_median_vals_df))
rows_d = dist(median_vals_pca$rotation[,1:6])
rowclust = hclust(as.dist(rows_d), "ward.D")


cols_d = dist(t(all_median_vals_df))

colclust = hclust(as.dist(cols_d), "ward.D")

library(pheatmap)

p = pheatmap(all_median_vals_df, 
             cluster_rows=rowclust,
             cluster_cols=colclust,
             )
```



For each cluster, plot the distribution of scaled features.

```{r}

for (cluster in sort(clusters)) {
 
   cluster_row_indices = which(fusion_data$leiden == cluster)
    
   feature_atts = as_tibble(data.scaled[cluster_row_indices,])
  
   ploc = feature_atts %>% gather() %>% ggplot(aes(x=key, y=value)) + 
     geom_boxplot() + 
     theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, size=rel(.9) )) +
     ggtitle(paste0("cluster: ", cluster))
   
   plot(ploc)
   
}



```



plot cosmic primary isoform fusion counts according to cluster

```{r}

cosmic_counts_per_cluster = fusion_data %>% filter(cosmic & is_primary) %>%
    group_by(fusion_name, leiden) %>% 
    tally()

cosmic_counts_per_cluster$leiden = factor(cosmic_counts_per_cluster$leiden, levels=rev(p$tree_row$labels[p$tree_row$order]))

empty_counts_df = data.frame(fusion_name="NA", leiden=factor(sort(p$tree_row$labels), levels=rev(p$tree_row$labels[p$tree_row$order])), n=0)

cosmic_counts_per_cluster = bind_rows(empty_counts_df, cosmic_counts_per_cluster)

# plot it

cosmic_sum_counts_per_cluster = cosmic_counts_per_cluster %>% 
    group_by(leiden) %>% summarize(sum_n = sum(n))

cosmic_all_iso_counts_per_cluster_plot = cosmic_counts_per_cluster %>% 
    ggplot(aes(x=leiden, y=n, fill=fusion_name)) + 
    geom_col() + 
    coord_flip() + 
    ggtitle("COSMIC primary isoform count")



plot(cosmic_all_iso_counts_per_cluster_plot)

```


# all fusions, top isoforms

```{r}
top_fusion_counts_per_cluster = fusion_data %>% 
    filter(is_primary) %>%
    select(fusion_name, leiden) %>% 
    group_by(fusion_name, leiden) %>% 
    tally()

top_fusion_counts_per_cluster$leiden = factor(top_fusion_counts_per_cluster$leiden, levels=rev(p$tree_row$labels[p$tree_row$order]))

empty_counts_df = data.frame(fusion_name="NA", leiden=factor(sort(p$tree_row$labels), levels=rev(p$tree_row$labels[p$tree_row$order])), n=0)

top_fusion_counts_per_cluster = bind_rows(empty_counts_df, top_fusion_counts_per_cluster)

sum_fusion_counts_per_cluster = top_fusion_counts_per_cluster %>% group_by(leiden) %>% summarize(sum_n = sum(n))

sum_fusion_counts_per_cluster_plot = sum_fusion_counts_per_cluster %>% ggplot(aes(x=leiden, y=sum_n)) + geom_col() + coord_flip() + ggtitle("Fusion primary isoform count")

plot(sum_fusion_counts_per_cluster_plot)
```

plot cosmic enrichment per cluster

```{r}

cosmic_sum_counts_per_cluster$fusion_type = "cosmic"
sum_fusion_counts_per_cluster$fusion_type = "all"
cosmic_and_all_counts_df = bind_rows(cosmic_sum_counts_per_cluster, sum_fusion_counts_per_cluster)

cosmic_and_all_counts_df = cosmic_and_all_counts_df %>% spread(key=fusion_type, value=sum_n)
cosmic_and_all_counts_df = cosmic_and_all_counts_df %>% mutate(frac_cosmic = cosmic/all)

cosmic_frac_enrichment_plot = cosmic_and_all_counts_df %>% ggplot(aes(x=leiden, y=frac_cosmic)) + 
    geom_col() + coord_flip() + ggtitle("Fraction COSMIC fusions")
plot(cosmic_frac_enrichment_plot)
```

Create single summary plot

```{r}


cosmic_all_iso_counts_per_cluster_plot_NLeg = cosmic_all_iso_counts_per_cluster_plot + 
    theme(legend.position = "none")

pg = plot_grid(p[[4]], cosmic_all_iso_counts_per_cluster_plot_NLeg, sum_fusion_counts_per_cluster_plot, cosmic_frac_enrichment_plot, 
          nrow=1, rel_widths = c(3,1,1,1) )

plot(pg)

LEIDEN_RESOLUTION=3

summary_png_filename = paste0("test.cluster_heatmap_n_counts.leiden_res", LEIDEN_RESOLUTION, ".png")
ggsave(pg, file=summary_png_filename, width=11)
summary_pdf_filename = paste0("test.cluster_heatmap_n_counts.leiden_res", LEIDEN_RESOLUTION, ".pdf")
ggsave(pg, file=summary_pdf_filename, width=11)

```

# Define cosmic-like clusters and likely artifact fusions, and annotate fusions accordingly
```{r}

# define the cosmic-like clusters for annotation purposes.
# here, we'll select those clusters that have at least 2 cosmic fusions

cosmic_like_clusters = fusion_data %>% 
    filter(is_primary & cosmic) %>% 
    group_by(leiden) %>% 
    tally() %>% 
    filter(n>=2) %>% 
    arrange(desc(n))

cosmic_like_clusters
```

```{r}
# annotate

# cosmic-peak-enriched cluster 4
# rest cosmic like clusters contain at least 2 cosmic fusions as per above.
cosmic_like_clusters = c(4, 6, 39, 38, 23, 8, 54, 13, 46, 34, 21, 7, 26, 10, 12, 3, 18)
pheatmap(all_median_vals_df[rownames(all_median_vals_df) %in% cosmic_like_clusters,],
         cluster_cols=colclust,
         main="cosmic-like"   
             )

# expr_microH_RT_artifact_clusters
expr_microH_RT_artifact_clusters =  c(47,20,41,15)
pheatmap(all_median_vals_df[rownames(all_median_vals_df) %in% expr_microH_RT_artifact_clusters,],
        cluster_cols=colclust,
        main="likely expr microH RT artifacts"        
             )



# high_FAR_microH_bioinf_artifact clusters
high_FAR_microH_bioinf_artifact_clusters = c(57,56,60)
pheatmap(all_median_vals_df[rownames(all_median_vals_df) %in% high_FAR_microH_bioinf_artifact_clusters,],
        cluster_cols=colclust,
        main="high FAR and microH, likely bioinf artifacts" 
             )

# high counter-evidence clusters
high_counter_evidence_clusters = c(49,51)
pheatmap(all_median_vals_df[rownames(all_median_vals_df) %in% high_counter_evidence_clusters,],
         cluster_cols=colclust,
         main="high counter-evidence"
             )


source("../Rlib/annotate_leiden_clusters.R")

fusion_data = annotate_leiden_clusters(fusion_data)

```

```{r}
# General statistics on cluster annotation types.

fusion_data %>% filter(is_primary) %>% group_by(fusion_cluster_att) %>% tally() %>% mutate(pct=prop.table(n))
```


```{r}
# define fusions enriched in "expr_microH_RT_artifact?" category.

fusions_expr_microH_RT_artifact_enriched = fusion_data %>% filter(is_primary) %>% group_by(fusion_name, fusion_cluster_att) %>% tally() %>% mutate(pct=prop.table(n)) %>% filter(fusion_cluster_att == "expr_microH_RT_artifact?") %>% filter(pct >= 0.5) 

fusions_expr_microH_RT_artifact_enriched 

```

# Examine feature distributions for certain fusions

```{r}

plot_fusion_scaled_atts = function(fusion_name) {

   row_indices = which(fusion_data$fusion_name == fusion_name & fusion_data$is_primary)
    
   feature_atts = as_tibble(data.scaled[row_indices,])
  
   p = feature_atts %>% gather() %>% ggplot(aes(x=key, y=value)) + 
     geom_boxplot() + 
     theme(axis.text.x = element_text(angle = 90, hjust = 1, vjust = 0.5, size=rel(.9) )) +
     ggtitle(paste0("fusion: ", fusion_name))
   
   return(p)
   
}

```

```{r}

plot_fusion_scaled_atts("TMPRSS2--ERG")

plot_fusion_scaled_atts("SS18--SSX1")
plot_fusion_scaled_atts("SS18--SSX2")

```



# Examine Left FAR and Right FAR for COSMIC fusions found in C4

```{r}

cosmic_fusions_in_C4 = fusion_data %>% filter(leiden == 4 & is_primary & cosmic) %>% select(fusion_name) %>% unique() %>% pull(fusion_name)

cosmic_FAR_data = fusion_data %>% filter(fusion_name %in% cosmic_fusions_in_C4) %>% filter(leiden==4 & is_primary) %>% 
  rename("5p-FAR" = FAR_left, "3p-FAR" = FAR_right ) %>%
  gather(key=FAR_type, value=FAR_val, "5p-FAR", "3p-FAR") %>%
  mutate(FAR_type=factor(FAR_type, levels=c("5p-FAR", "3p-FAR"))) 

cosmic_FAR_data %>%
  ggplot(aes(x=FAR_type, y=FAR_val)) + geom_boxplot() + facet_wrap(~fusion_name, scale='free_y')
```

- used for Figure S5:

```{r}

cosmic_FAR_data %>%
  ggplot(aes(x=FAR_type, y=FAR_val, fill=FAR_type)) + 
  geom_violin(scale='width') +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(1))) + 
  facet_wrap(~fusion_name, scale='free_y') +  
  geom_jitter(size=rel(0.1)) +
  theme(strip.text.x = element_text(size = rel(0.8)))

```


```{r}

cosmic_paired_FAR_data = cosmic_FAR_data %>% select(sample_name, fusion_name, FAR_type, FAR_val) %>% spread(key=FAR_type, value=FAR_val)

cosmic_fusions_here = cosmic_paired_FAR_data %>% select(fusion_name) %>% unique() %>% pull(fusion_name)

cosmic_FAR_ttest_pvals = NULL

for (cosmic_fusion_select in cosmic_fusions_here) {
  
  paired_FAR_data = cosmic_paired_FAR_data %>% filter(fusion_name == cosmic_fusion_select)
  
  left_FAR_vals = paired_FAR_data %>% pull(`5p-FAR`)
  right_FAR_vals = paired_FAR_data %>% pull(`3p-FAR`)
  
  if (length(left_FAR_vals) > 1) {
    t = t.test(left_FAR_vals, right_FAR_vals, alternative = 'less', paired = T)
    cosmic_FAR_ttest_pvals = bind_rows(cosmic_FAR_ttest_pvals, data.frame(fusion=cosmic_fusion_select, tp_val = t$p.value))
  }
}


```

```{r}

cosmic_FAR_ttest_pvals$tp_val_BH = p.adjust(cosmic_FAR_ttest_pvals$tp_val, method = 'BH')

cosmic_FAR_ttest_pvals = cosmic_FAR_ttest_pvals %>% mutate(signif = tp_val_BH < 0.05)


cosmic_FAR_ttest_pvals %>% mutate(negLogP = -1*log10(tp_val_BH)) %>%
  ggplot(aes(x=fusion, y=negLogP, fill=signif)) + geom_col() + 
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1, size=rel(1))) 

```




```{r}
run_fishers_exact_test = function(yes_cosmic, no_cosmic, tot_cosmic_fusions, tot_noncosmic_fusions) {
    
    confusion_matrix = matrix(c(yes_cosmic, no_cosmic,
                                tot_cosmic_fusions - yes_cosmic, tot_noncosmic_fusions - no_cosmic),
                              nrow=2, byrow = T)    
    
    f = fisher.test(confusion_matrix, alternative = 'greater')
    
    return(f$p.value)
    
}


cosmic_cluster_count_info = fusion_data %>% filter(is_primary) %>%
    group_by(leiden, cosmic) %>% tally() %>% 
    spread(cosmic, n, fill=0) %>% rename(yes_cosmic = `TRUE`, no_cosmic = `FALSE`)

tot_cosmic_fusions = sum(cosmic_cluster_count_info$yes_cosmic)
tot_noncosmic_fusions = sum(cosmic_cluster_count_info$no_cosmic)

cosmic_cluster_info = cosmic_cluster_count_info %>% mutate(pval = 
                                                               run_fishers_exact_test(
                                                                   yes_cosmic, no_cosmic, 
                                                                   tot_cosmic_fusions, tot_noncosmic_fusions)
)


cosmic_cluster_info = cosmic_cluster_info %>% mutate(negLOGp = -1 * log10(pval))



cosmic_cluster_info %>% ggplot(aes(x=leiden, y=negLOGp)) + geom_col() 


```


```{r}


cosmic_cluster_info %>% ggplot(aes(x=no_cosmic+yes_cosmic, y=yes_cosmic, color = negLOGp)) + geom_point() 


```


```{r}
cosmic_cluster_info %>% ungroup() %>%  mutate(tot_fusions = yes_cosmic + no_cosmic) %>%
    mutate(pct_all_cosmic_fusions=prop.table(yes_cosmic))  %>%
    mutate(pct_tot_fusions = prop.table(tot_fusions)) %>% 
               arrange(desc(negLOGp))


```
```{r}
# find those cosmic clusters having at least two instances. 
#  we'll call all of these the cosmic-like fusion clusters

cosmic_cluster_info %>% filter(yes_cosmic >= 2)

```


```{r}
# count gene pairs per cluster


fusion_data %>% filter(is_primary) %>% group_by(leiden) %>% tally()

fusion_data %>% filter(is_primary) %>% select(leiden, fusion_name) %>% unique() %>% group_by(leiden) %>% tally()


```

